\chapter{Methodology / project development}
\label{chap:methodology}

This chapter is structured as the route that would be followed if the project
were started again with the final evidence already known. The detailed
chronology is preserved in Appendix~A, but the main methodology now follows the
fastest defensible path: fix the evaluation contract, use the early neural
models as diagnostics, identify the target distributions, calibrate frozen
priors, and train a bounded residual model around those priors. The emphasis is
on decisions that affect reproducibility: masking, splitting, conditioning, and
the distinction between raw deep prediction and prior-anchored residual
prediction.

\section{Dataset protocol}

Each sample is a complete $513\times513$ pixel CKM. The spatial resolution is
\SI{1}{m}\(\times\)\SI{1}{m} per pixel, so each map covers a
\SI{513}{m}\(\times\)\SI{513}{m} square. The UAV transmitter is horizontally
fixed at the centre of the map, while its antenna height \(h_{\mathrm{tx}}\)
varies from sample to sample and is treated as a primary conditioning variable.
Every non-building pixel is an active receiver location, with the receiver antenna placed exactly at
\SI{1.5}{m} above ground. Building pixels are not receiver positions and are
excluded from the training loss and reported metrics rather than being counted
as easy background pixels. The model never receives a random crop as an
independent training city. Instead, the split is performed at city level. For
the distribution-first and final neural models the split uses
\texttt{split\_seed = 42}, with 15\% validation and 15\% test samples after
city assignment. This avoids the most common optimistic error in spatial radio
prediction: training on one part of a city and testing on another visually
similar part of the same city.

The transmitter-height samples are not uniformly distributed. Comparing the
stored HDF5 histogram with the tested affine beta-mixture variants shows that
the closest nominal sampler is
\(h_{\mathrm{design}}\approx 10+490X\), where
\[
X\sim 0.75\,\mathrm{Beta}(3,23)+0.25\,\mathrm{Beta}(4,4).
\]
However, the finite HDF5 database is the relevant distribution for training
and reporting. Reading the stored \texttt{uav\_height} field for all 16180
samples gives an empirical range of \SIrange{11.61}{477.54}{m}, usually
rounded in the text as about \SIrange{12}{478}{m}; \SI{477.54}{m} is the
largest sampled HDF5 value, not a hard endpoint of the approximate beta law.
This is statistically normal: under \(10+490X\),
\(P(h_{\mathrm{tx}}>\SI{478}{m})=3.19\times10^{-5}\) per draw, so only
0.52 such samples are expected in 16180 draws and the probability of drawing
none is about 0.60. A reproducible formulaic version of the realized database
distribution is the \SI{3}{m} empirical density. Let \(\Delta h=\SI{3}{m}\),
\(B_k=[10+3k,\,13+3k)\), and let \(h_i\) be the stored height of
sample \(i\). Then
\begin{equation}
  \begin{aligned}
  \hat f_{\Delta h}(h) &= d_k,\quad h\in B_k,\\
  d_k &= \frac{n_k}{N\Delta h},\quad
  n_k=\sum_{i=1}^{N}\mathbf{1}\{h_i\in B_k\},\quad N=16180 .
  \end{aligned}
  \label{eq:height_empirical_histogram}
\end{equation}
When the density is visualized, the plotted curve linearly interpolates
between the bin-centre values \((c_k,d_k)\), with \(c_k=11.5+3k\).
The all-database height counts are 3972 samples below \SI{50}{m}, 8456 from
\SIrange{50}{150}{m}, 2490 from \SIrange{150}{300}{m}, and 1262 above
\SI{300}{m}. In the actual training cities, the same bins contain 2676, 5654,
1644, and only 866 samples, respectively. This matters for interpreting the
height-binned results: the higher-altitude bins contain fewer training
examples, so their lower RMSE later in the thesis is evidence that those cases
are physically easier, not that they are better represented in the training
set.

The prior artifacts use the same city-holdout contract as the final model:
\texttt{val\_ratio = 0.15}, \texttt{test\_ratio = 0.15}, and
\texttt{split\_seed = 42}. The Try~78 path-loss prior and Try~79 spread priors
are fitted on the 10840 training samples. Standalone prior audits can then use
the full held-out validation+test pool of 5340 samples, while final
model-versus-prior comparisons use the 2590-sample final test split. This keeps
training cities, validation cities, and final test cities separated throughout
the prior and residual-model evaluations.

\begin{table}[ht]
\centering
\scriptsize
\caption{City-holdout assignment used by the distribution-first and final neural models.}
\label{tab:split_city_assignment}
\begin{tabularx}{\textwidth}{lrrX}
\toprule
Split & Samples & Cities & City names \\
\midrule
Train & 10840 & 37 & Abidjan, Abu Dhabi, Alexandria, Antigua Guatemala,
Athens, Bangalore, Bangkok, Bath, Berlin, Bratislava, Brugge, Buenos Aires,
Canberra, Chefchaouen, Dalian, Hanoi, Ho Chi Minh City, Hyderabad, Kinshasa,
Leuven, Ljubljana, Luanda, Lyon, Manaus, Minsk, Nagpur, Paris,
Rio de Janeiro, Santiago, Shanghai, Siem Reap, St.~Louis, Thane, Toledo,
Valparaiso, Venice, Vienna. \\
Validation & 2750 & 15 & Bilbao, Chiang Mai, Cleveland, Dhaka, Dubrovnik,
Fortaleza, Marrakesh, Mexico City, Munich, Nazareth, Nice, Salvador, Segovia,
Tunis, Victoria. \\
Test & 2590 & 14 & Barcelona, Beijing, Carcassonne, Copenhagen, Halifax,
Jaipur, Johannesburg, Key West, Kuala Lumpur, Osaka, Phuket, Pune, Surat,
Vancouver. \\
\bottomrule
\end{tabularx}
\normalsize
\end{table}

The receiver mask is:
\begin{equation}
  m_{\mathrm{ground}}(i,j) =
  \begin{cases}
    1, & \text{if } \mathrm{topology}(i,j)=0
    \text{ (non-building receiver pixel)},\\
    0, & \text{otherwise.}
  \end{cases}
  \label{eq:ground_mask}
\end{equation}
Equivalently, the CKM supervision grid contains one receiver at every
non-building pixel, all at the fixed receiver height
\(h_{\mathrm{rx}}=\SI{1.5}{m}\). The mask only removes building pixels from
the loss and metrics; it does not subsample the remaining ground pixels.
The LoS/NLoS visibility masks used later by the priors and the final
residual model are geometric support inputs rather than target radio maps. In
the CKM experiments they are read from the dataset when available; for
topology-only deployment they can be generated from topology and UAV antenna
height using geometric ray-casting, although generated masks are not expected
to be bit-identical to the CKM reference masks. This makes the visibility
signal explicit without requiring a previous prediction of the target radio
maps.
All RMSE values in this thesis are pixel-weighted over valid ground pixels:
\begin{equation}
  \mathrm{RMSE}
  =
  \sqrt{
  \frac{\sum_{s,i,j} m_s(i,j)\left(\hat y_s(i,j)-y_s(i,j)\right)^2}
       {\sum_{s,i,j} m_s(i,j)}
  }.
  \label{eq:rmse}
\end{equation}
This is stricter than averaging per-sample RMSEs because large valid maps
contribute proportionally to the number of receiver pixels.

\section{Diagnostic baselines retained from development}

The first implementation used a U-Net generator with a PatchGAN discriminator.
The generator predicted channel maps from topology-related inputs, while the
discriminator encouraged realistic local patches. This was useful as a
software prototype but not as a final scientific direction. The adversarial
term did not improve the physical errors that dominated validation RMSE, and it
made diagnosis harder.

The next phase focused on controlled engineering changes. Transposed
convolutions were replaced by bilinear upsampling and projection to reduce
checkerboard artifacts. Multi-scale supervision was added so that the model
could learn both global radial structure and local map details. The combined
decoder/multiscale run became
the strongest clean pre-prior path-loss baseline in this family.

After the masking correction, building pixels were excluded from both training loss and
validation metrics. This changed the interpretation of earlier numbers: any
method that implicitly benefited from easy building pixels was no longer
acceptable. The ground-only rule became an invariant for later models.

\section{Methodology structure after the development lessons}

The work was not a single clean jump from U-Net to the final residual GMM-head
model. The detailed numbering is kept in Appendix~A; the methodology only
retains the lessons that affect the final design. If the project were repeated
from the current endpoint, the efficient route would be:

\begin{enumerate}
\item define the city-holdout split and ground-only metric before comparing
      any architecture;
\item use simple U-Net and cGAN baselines only to verify the data pipeline and
      expose why visual realism is not enough;
\item separate LoS/NLoS, topology, and UAV-height regimes before interpreting
      the errors;
\item diagnose target histograms before choosing the output head;
\item calibrate the path-loss and spread priors as train-only estimators;
\item for path loss, reproduce the final prior-anchored residual GMM-head setup
      rather than restarting another raw backbone search;
\item for delay and angular spread, add an explicit high-tail or outlier
      diagnostic before final fine-tuning, because the remaining errors are
      concentrated in rare spread events;
\item train the final network as a bounded residual corrector over the frozen
      priors rather than as a raw map generator.
\end{enumerate}

The earlier models are therefore not deleted from the thesis narrative. They
remain as controlled evidence for why the final method is organised around the
evaluation contract, distribution shape, and prior-anchored residual learning
instead of around a larger generic backbone. This is a change in exposition,
not a change in the experimental contract: the dataset, the city-holdout
split, the ground-only mask, the final prior-anchored architecture, and the detailed
path-loss-prior, spread-prior, and residual-GMM formulas remain the same. The older version of the methodology
read more like a chronological development diary; this reduced version keeps
the same evidence but presents it as the design path that now looks defensible
after the negative results are known. The chronology still matters because it
explains why each simplifying decision was made: U-Net and cGAN runs validated
the pipeline, PMNet and PMHHNet diagnosed the limits of stronger backbones,
the distribution-first attempts exposed the target marginals, and the final
prior-anchored model combined the surviving pieces.

\section{Physical priors and PMHHNet}

The middle development phase moved from direct map prediction to residual learning, as
formulated in Eq.~\ref{eq:prior_residual}. This phase was heavily influenced by
the \texttt{PMNet} architecture \cite{pmnet2023}, which established the
viability of using U-Net-like backbones for path-loss map prediction in
terrestrial cellular networks.
\begin{equation}
  \hat{\textbf{Y}}
  =
  \textbf{P}_{\mathrm{phys}}
  +
  f_{\theta}\!\left(\mathcal{T}, h, \textbf{P}_{\mathrm{phys}}\right),
  \label{eq:prior_residual}
\end{equation}
where \(\textbf{P}_{\mathrm{phys}}\) is the analytical or calibrated prior map
and \(f_{\theta}\) predicts only the residual correction that the prior cannot
explain.

\subsection{Mathematical Evolution: From PMNet to PMHHNet}

The \texttt{PMHHNet} (Path-loss Map with High-frequency and Height-awareness
Network) was developed as a specific refinement of the \texttt{PMNet}
philosophy tailored for the UAV-enabled Air-to-Ground (A2G) scenario.

\paragraph{High-Frequency (HF) Stem.}
Standard convolutional encoders tend to smooth out sharp building transitions
during downsampling. To counter this, \texttt{PMHHNet} introduces a dedicated
high-frequency stem. Let $\mathcal{T}$ be the topology map. We compute the
discrete Laplacian $|\nabla^2 \mathcal{T}|$ to identify building edges and
shadow boundaries. The initial feature map $\textbf{X}_0$ is computed as:
\begin{equation}
  \textbf{X}_0 = \mathrm{Stem}(\mathcal{T}) + \mathrm{Proj}(|\nabla^2 \mathcal{T}|)
  \label{eq:hf_stem}
\end{equation}
where $\mathrm{Proj}$ is a $1\times1$ convolution that maps the edge-detector
output to the stem's feature depth. This ensures that building geometry is
strongly injected into the earliest layers of the encoder.

\paragraph{Continuous Sinusoidal FiLM Conditioning.}
While \texttt{PMNet} was designed for fixed base-station heights, UAVs require
continuous height generalization. \texttt{PMHHNet} utilizes Feature-wise Linear
Modulation (FiLM) \cite{perez2018film} to inject the scalar height $h$.
The height is first mapped to a sinusoidal embedding $\textbf{e}(h) \in \mathbb{R}^K$
(Eq.~\ref{eq:sin_film}). For any feature map $\textbf{X} \in \mathbb{R}^{C \times H \times W}$
at depth $l$, the conditioned map $\textbf{X}'$ is:
\begin{equation}
  \textbf{e}(h)
  =
  \left[
    \sin(\omega_1 \tilde{h}), \cos(\omega_1 \tilde{h}), \ldots,
    \sin(\omega_{K/2} \tilde{h}), \cos(\omega_{K/2} \tilde{h})
  \right],
  \qquad
  \tilde{h} = \frac{h-h_{\min}}{h_{\max}-h_{\min}},
  \label{eq:sin_film}
\end{equation}
\begin{equation}
  \textbf{X}' = \textbf{X} \odot (1 + \textbf{g}_l(\textbf{e}(h))) + \textbf{b}_l(\textbf{e}(h))
  \label{eq:film_math}
\end{equation}
where $\textbf{g}_l$ and $\textbf{b}_l$ are learned linear projections mapping the
embedding to the channel dimension $C$. This pattern, inspired by time-step
conditioning in diffusion models \cite{dhariwal2021diffusion}, allows the
network to modulate its internal representation dynamically based on the UAV's
altitude.

\subsection{Limitations and Legacy}

Despite these improvements, the early residual phase plateaued at about
\SI{9.4}{dB}. The primary bottleneck was the "soft" nature of the
residual learning; the network often under-compensated for the physical
propagation law in NLoS regions.

\begin{quote}
\textit{Legacy of PMHHNet:} While the full architecture was eventually superseded
by the final prior-anchored GMM approach, its core components, the
HF stem and sinusoidal FiLM, remain highly effective for point-prediction
tasks or real-time implementations where the distributional complexity of a GMM
is not required.
\end{quote}

\section{Distribution-first modelling}
\label{sec:dist_first}

The decisive late-stage diagnosis was that NLoS path loss was not merely a
harder regression target; it possessed a fundamentally different error
distribution.  This diagnosis came from comparing predicted and ground-truth
histograms: the maps showed where the model was wrong, but the histograms
showed \emph{what} the model had misunderstood.  The central idea of the late
pipeline was therefore to identify the target distribution before asking a
network to perform pixel-by-pixel prediction. The distribution-first path-loss
model addressed this by
emitting a Gaussian Mixture Model (GMM)-parameterized residual head rather than
only a direct scalar map.

\begin{figure}[htbp]
\centering
\resizebox{0.9\linewidth}{!}{%
\begin{tikzpicture}[
    >=stealth, thick,
    nd/.style={draw, rectangle, rounded corners=3pt, minimum width=2.6cm,
               minimum height=0.65cm, align=center, font=\scriptsize},
    inp/.style={nd, fill=blue!10},
    sa/.style={nd, fill=red!10},
    sb/.style={nd, fill=green!10},
    ob/.style={nd, fill=orange!10},
    grp/.style={draw=orange!50, fill=yellow!8, rounded corners=4pt, inner sep=8pt},
    lbl/.style={font=\tiny\itshape, anchor=south west, yshift=2pt},
]

%% Inputs row
\node[inp] (topo)  {topology\_map / 90};
\node[inp, right=0.4cm of topo]  (masks) {los, nlos,\\ground masks};
\node[inp, right=0.4cm of masks] (hgt)   {UAV height\\embedding};
\node[nd,  right=1.4cm of hgt]   (filt)  {Invalid\\target filter};

%% Routing row
\node[nd, fill=blue!15, below=0.5cm of topo]             (cls) {6 topology classes};
\node[nd, fill=blue!20, minimum width=3.2cm,
      below=0.5cm of masks, xshift=0.6cm]                (exp) {12 experts = 6 classes $\times$\\LoS / NLoS};
\node[nd, fill=blue!15, below=0.5cm of filt]             (vld) {valid\_target\\finite and $\ge 0$};

\draw[->] (topo) -- (cls);
\draw[->] (cls)  -- (exp);
\draw[->] (masks)-- (exp);
\draw[->] (hgt)  -- (exp);
\draw[->] (filt) -- (vld);

\begin{pgfonlayer}{background}
  \node[grp, fit=(topo)(masks)(hgt)(filt)(cls)(exp)(vld)] (inp_box) {};
\end{pgfonlayer}
\node[lbl] at (inp_box.north west) {Inputs and routing};

%% Stage A: GMM head (right)
\node[sa, below=1.2cm of exp]       (enc)  {Encoder with FiLM(height)};
\node[sa, below=0.4cm of enc]       (gap)  {Global average pooling};
\node[sa, below=0.4cm of gap]       (gmm)  {GMM head};
\node[sa, below=0.4cm of gmm]       (outa) {$\pi_k,\,\mu_k,\,\sigma_k$};

\draw[->] (exp) -- (enc);
\draw[->] (enc) -- (gap);
\draw[->] (gap) -- (gmm);
\draw[->] (gmm) -- (outa);

\begin{pgfonlayer}{background}
  \node[grp, fit=(enc)(gap)(gmm)(outa)] (sa_box) {};
\end{pgfonlayer}
\node[lbl] at (sa_box.north west) {Stage A: distribution head};

%% Stage B: map head (left of Stage A)
\node[sb, left=1.6cm of enc, yshift=-0.4cm]  (dec)  {U-Net decoder};
\node[sb, below=0.4cm of dec]                 (cond) {Condition on height +\\GMM summary};
\node[sb, below=0.4cm of cond]                (pix)  {Per-pixel $p_k,\,z_k,\,\log\sigma_k$};
\node[sb, below=0.4cm of pix]                 (yhat) {$\hat{y}=\sum p_k(\mu_k+z_k\sigma_k)$};

\draw[->] (dec)  -- (cond);
\draw[->] (cond) -- (pix);
\draw[->] (pix)  -- (yhat);
\draw[->] (outa.west) -- ++(-0.25,0) |- (cond.east);

\begin{pgfonlayer}{background}
  \node[grp, fit=(dec)(cond)(pix)(yhat)] (sb_box) {};
\end{pgfonlayer}
\node[lbl] at (sb_box.north west) {Stage B: map head};

%% Objective row
\node[ob, below=1.2cm of yhat]     (o1) {map\_nll under\\mixture};
\node[ob, right=0.3cm of o1]       (o2) {moment\_match};
\node[ob, right=0.3cm of o2]       (o3) {rmse\_db};
\node[ob, right=0.3cm of o3]       (o4) {outlier\_budget};
\node[ob, right=1.0cm of o4]       (o5) {dist\_kl on\\empirical hist.};

\draw[->] (yhat.south) -- ++(0,-0.4) -| (o1.north);
\draw[->] (yhat.south) -- ++(0,-0.4) -| (o2.north);
\draw[->] (yhat.south) -- ++(0,-0.4) -| (o3.north);
\draw[->] (yhat.south) -- ++(0,-0.4) -| (o4.north);
\draw[->] (outa.south) -- ++(0,-0.4) -| (o5.north);

\begin{pgfonlayer}{background}
  \node[grp, fit=(o1)(o2)(o3)(o4)(o5)] (ob_box) {};
\end{pgfonlayer}
\node[lbl] at (ob_box.north west) {Training objective on valid/bound pixels};

\end{tikzpicture}%
}
\caption[Distribution-first path-loss system. Stage A learns global GMM
parameters from pooled features and Stage B predicts per-pixel assignments.]%
{Distribution-first path-loss system. Stage~A learns global
GMM parameters $(\pi_k,\mu_k,\sigma_k)$ from pooled features conditioned on
UAV height; Stage~B produces per-pixel soft assignments and
within-component offsets anchored to those parameters.}
\label{fig:try76_system}
\end{figure}

The distribution-first spread model applied the same two-stage idea to delay
spread and angular spread, but changed the mixture shape to match the targets:
a near-zero LoS spike plus a heavy NLoS tail. The architecture is therefore not
redrawn as a second full diagram; it follows the same routing and Stage-A /
Stage-B structure as Figure~\ref{fig:try76_system}, with the path-loss LoS/NLoS
experts replaced by delay-spread and angular-spread experts.

For path loss, the distribution-first model utilized 12 specialized experts: six topology classes partitioned by LoS
and NLoS regions. As shown in Figure~\ref{fig:try76_system}, Stage A acts as a distribution head, predicting the global GMM parameters
$(\pi_k,\mu_k,\sigma_k)$ from map features and altitude conditioning.
Stage B then decodes these priors spatially, predicting per-pixel soft component assignments and within-component
offsets. The final map reconstruction can be written as:
\begin{equation}
  \hat y(i,j)
  =
  \sum_{k=1}^{K} p_k(i,j)
  \left(\mu_k + z_k(i,j)\sigma_k\right).
  \label{eq:gmm_reconstruction}
\end{equation}

The spread counterpart extended this distribution-first philosophy to delay spread and angular spread. These targets
are not cleanly Gaussian; their empirical distributions resemble a near-zero spike (representing the LoS direct path) combined with a
heavy tail (representing NLoS scattering). Standard Mean Squared Error (MSE) objectives tend to underrepresent these rare
but physically significant heavy-tail spread values, necessitating a dedicated mixture head.

\subsection{How to construct a good deep-learning model for radio maps}
\label{sec:good_dl_model_design}

The main lesson from the deep-learning experiments is that architecture is
secondary to the question: \emph{what statistical object is the network being
asked to predict?} If the project were restarted with the final evidence
available, the first step would be to define the city-holdout, ground-only
metric contract; the second would be to inspect the target distributions; and
only then would the backbone, loss, prior, and routing design be chosen. The
project did not discover this in one step. It moved from direct image
translation, to residual learning over physical estimates, to topology experts,
to distribution-first heads, and finally to frozen priors plus bounded neural
residuals.

This evolution reframes the role of priors. Midway through the project, the
best models were residual-over-prior systems, which made the prior appear to be
the essential ingredient. The distribution-first path-loss model changed that
interpretation. It used no COST-231 prior, no enhanced two-ray prior, no
knife-edge channel, and no residual-over-prior decomposition, yet it reached
NLoS results close to the final prior family by modelling the NLoS marginal
directly. Therefore, the prior is useful but not magic: it is not external
knowledge that solves the task by itself. It is a stable and auditable estimator
of large-scale structure, especially when that structure is simpler to fit than
to relearn with a high-capacity network.

The final pipeline reintroduces priors without contradicting the
distribution-first lesson. The histogram diagnosis showed that the PMHHNet
family was failing because the output distribution was wrong, not simply
because one more physical channel was missing. The first response was therefore
to fix the deep-learning side and prove that NLoS path loss can be modelled
without a prior when the head represents the target distribution. Once that was
clear, the question became practical: which parts of the prediction are
cheapest, most stable, and easiest to audit? For LoS path loss, a calibrated
two-ray/radial model is more efficient than asking a CNN to relearn the same
ring structure. For blocked NLoS path loss and the spread targets, the neural
model is still valuable because the problem is no longer just estimating a
smooth mean; it is placing local residuals and rare tails in the right parts of
the map.

Path loss also cannot be treated as one homogeneous target. LoS is not only the
easy subset: it contains a deterministic spatial law driven by distance,
height, and coherent two-ray structure. A convolutional network can learn a
smooth radial trend from distance and topology channels, but it does not
naturally learn the calibrated ring pattern if the objective rewards only the
average pixel error. NLoS is different. In valid CKM ground pixels its marginal
is concentrated around a blocked-link attenuation level, but it is noisy,
topology-dependent, and hard to place spatially. The model must answer two
questions at once: which NLoS value range belongs to this map, and which pixels
belong to the high-error tail. The old PMHHNet family exposed this mismatch
through severe NLoS mode collapse, with predicted NLoS means around
\SI{36.5}{dB} against target means around \SI{107.1}{dB}.

Delay spread and angular spread are even more distributional. In LoS regions,
the direct path creates a near-zero spike: small delay spread and narrow angular
spread are physically expected. In NLoS regions, the spread targets often carry
the useful channel information because reflected, diffracted, and scattered
paths dominate the received energy. A high delay spread or wide angular spread
is therefore not merely an error-prone outlier; it is evidence that multipath
has become important. This is why a plain RMSE or Huber head tends to miss the
engineering-relevant tail: it can improve the common near-zero region while
underpredicting rare high-spread areas. The late histogram diagnostics that
motivated this shift are kept in Appendix~\ref{app:qualitative_diagnostics};
the methodology keeps their design consequence: spread maps need explicit
spike-plus-tail or high-tail diagnostics, not just smoother images.

The resulting rule set is compact but chronological. First, identify whether
the target is smooth, narrow, multi-modal, spike-plus-tail, or heavy-tailed
before choosing MSE, Huber, GMM, NLL, or residual losses. Second, treat LoS and
NLoS as semantic regimes, not only as reporting columns. Third, keep masks,
units, city holdout, topology class, antenna height, and tail-risk policy
visible during validation. Finally, use visual realism only as a qualitative
check: cGAN sharpness and larger PMNet/PMHHNet-style backbones did not solve
the physical RMSE problem until the target distribution itself was represented.

\section{GMM component count and distribution choice}
\label{sec:gmm_k_choice}

The Stage-A distribution heads in the distribution-first models
represent each per-sample target marginal as a parametric family. Choosing
the family and its number of components $K$ is not a free hyperparameter; it
is constrained by the empirical shape of the CKM histograms. A dedicated
distribution-ranking diagnostic fits a fixed shortlist of parametric models
to the aggregated validation histograms of every target under each topology
class and LoS/NLoS split, and ranks them by the KL divergence
\(\mathrm{KL}(p_{\mathrm{empirical}}\;||\;p_{\mathrm{model}})\).  The shortlist
contains Gaussian, Laplace, skew-Normal, log-Normal, Gamma, Weibull, a
spike-plus-exponential model for near-degenerate targets, and Gaussian
mixtures with $K\in\{2,3,4,5\}$.

Table~\ref{tab:gmm_k_choice} summarises the winning family for the four
targets used in this thesis, aggregated across all cities. The table uses the
target-side (ground-truth) histograms, because the distribution-head objective
is defined against the ground-truth marginal, not the current prediction.

\begin{table}[!htbp]
\centering
\caption[Best-fit parametric family by target.]%
{Best-fit parametric family for each target distribution on the CKM
validation histograms, across all cities. Based on the validation
KL-ranking study.}
\label{tab:gmm_k_choice}
\begin{tabularx}{\textwidth}{lcX}
\toprule
Target & Best family & Shape notes \\
\midrule
Path loss, LoS   & GMM-4 & Single dominant mode near \SI{97}{dB} with a mild
                              low-tail; GMM-4 wins on the raw target
                              histogram, but a Gaussian is already
                              competitive. \\
Path loss, NLoS  & GMM-5 & Narrow peak near \SI{107}{dB} plus a secondary
                              low-variance mode; a single Gaussian is too
                              wide. \\
Delay spread     & GMM-5 & Near-zero spike plus a long exponential tail;
                              the spike-plus-exponential family is a close
                              second. \\
Angular spread   & spike-plus-exp. & Dirac-like spike near 0\textdegree{} plus
                              a heavy tail; GMM can approximate it but
                              spike-plus-exp is the native family. \\
\bottomrule
\end{tabularx}
\end{table}

Three consequences follow. First, a single Gaussian (or any unimodal head) is
never the best family for NLoS path loss or for the two spreads. This is
exactly the regime in which the earlier PMHHNet family produced mode collapse
with a single-output MSE head. Second, the number of components that wins is
consistently $K\in\{4, 5\}$ for the raw path-loss and spread targets, not
$K=3$. The Stage-A heads in this thesis therefore use $K=5$ by default on
the target-distribution experts, even though the code supports
$K\in\{2,\dots,5\}$. The final \texttt{num\_components} field
configuration (\(K=3\)) controls a different quantity: the shared backbone's
residual distribution head after the frozen priors have already
explained the main LoS, NLoS, delay-spread, and angular-spread structure. In
that residual space, three modes are sufficient to model negative correction,
near-zero correction, and positive correction, while avoiding a larger mixture
that would be harder to stabilize during the final multi-task training. Third, the
angular-spread target is better served by a native spike-plus-exponential
head than by a GMM, because its marginal is effectively degenerate at
\ang{0}; this is implemented directly in the distribution-first angular-spread
experts.

\section{Loss functions of the final pipeline}
\label{sec:loss_functions}

The detailed definitions of the final objective are given in
Sec.~\ref{sec:try80_objective}. This section keeps only the training contract
needed to compare the model families. For each target
\(t\in\{\mathrm{PL},\mathrm{DS},\mathrm{AS}\}\), the implementation defines a
composite masked loss
\begin{equation}
\begin{aligned}
\mathcal{L}
&=
\sum_t w_t
\left(
\lambda_{\mathrm{NLL}}\mathcal{L}_{\mathrm{NLL},t}
+\lambda_{\mathrm{KL}}\mathcal{L}_{\mathrm{KL},t}
+\lambda_{\mathrm{mom}}\mathcal{L}_{\mathrm{mom},t}
\right.\\
&\hspace{2.5cm}\left.
+\lambda_{\mathrm{anchor}}\mathcal{L}_{\mathrm{anchor},t}
+\lambda_{\mathrm{guard}}\mathcal{L}_{\mathrm{guard},t}
+\lambda_{\mathrm{out}}\mathcal{L}_{\mathrm{out},t}
\right.\\
&\hspace{2.5cm}\left.
+\lambda_{\mathrm{RMSE}}\mathcal{L}_{\mathrm{RMSE},t}
+\lambda_{\mathrm{MAE}}\mathcal{L}_{\mathrm{MAE},t}
\right).
\end{aligned}
\label{eq:try80_total}
\end{equation}
All sums are restricted to valid receiver pixels.

\begin{table}[!htbp]
\centering
\caption{Active loss settings in the final residual GMM-head model stages.}
\label{tab:try80_loss_settings}
\scriptsize
\begin{tabularx}{\textwidth}{>{\raggedright\arraybackslash}p{0.22\textwidth}
>{\raggedright\arraybackslash}p{0.19\textwidth}
>{\raggedright\arraybackslash}p{0.19\textwidth}
>{\raggedright\arraybackslash}X}
\toprule
Term group & Large-capacity residual stage & Final residual fine-tuning stage & Interpretation \\
\midrule
NLL, KL, moment and outlier budget &
1.00 / 0.25 / 0.10 / 0.02 &
0 / 0 / 0 / 0 &
Active in the preceding large-capacity stage, then disabled in the selected
deterministic fine-tuning stage. \\
RMSE / MAE &
1.50 / 0.10 &
1.00 / 0.05 &
Supervised pixel losses in native output units; the final fine-tuning stage
makes them the dominant active terms. \\
Anchor / prior guard &
0.05 / 0.10 &
0.01 / 0.02 &
Keep residual corrections small and penalize model states that become worse
than the frozen priors, with lighter regularization in the fine-tuning stage. \\
Task weights &
PL 1.00, DS 1.00, AS 1.00 &
PL 1.00, DS 0.25, AS 0.25 &
Equal target weighting before the final fine-tuning stage; then path loss
becomes the main optimization target while spread supervision remains in the
joint model. \\
\bottomrule
\end{tabularx}
\end{table}

This is a deliberate simplification relative to the earlier branches. The
PMHHNet diagnostic runs used masked MSE with multi-scale and Laplacian-style auxiliaries; the
distribution-first models removed those structural losses and focused on target
shape. The final model keeps the residual GMM-head machinery available: it
still emits mixture parameters and uses the local-weighted model-space mixture
mean as the deterministic point prediction; for delay and angular spread this
mean is computed in \(\log(1+y)\) space and inverse-transformed before native
RMSE/MAE are reported. The probabilistic NLL, KL, moment and outlier
terms were active in the preceding large-capacity training stage, but the
selected final model is chosen after a deterministic RMSE/MAE-dominant
fine-tuning stage because the frozen priors already explain the main
large-scale structure. Consequently, the final claim is an RMSE improvement
over the frozen priors, not a calibrated probabilistic prediction claim.

\section{Training hyperparameters}
\label{sec:hyperparams}

Table~\ref{tab:hyperparams} summarises the training configuration of the
final prior-anchored pipeline. The values for the distribution-first predecessors
and the final residual GMM-head model are reported as methodological settings
rather than as file-path artifacts. Older runs are reported more compactly
because the final thesis line inherits only the architectural lessons from
them, not the training stack.

\begin{table}[!htbp]
\centering
\caption[Training configuration of the final pipeline.]%
{Training configuration of the final pipeline. Unless otherwise
stated, all runs use AdamW, image size $513\times513$, batch size 1 per GPU,
and gradient accumulation over 8 steps.}
\label{tab:hyperparams}
\scriptsize
\begin{tabularx}{\textwidth}{>{\raggedright\arraybackslash}p{0.17\textwidth}
                            >{\raggedright\arraybackslash}X
                            >{\raggedright\arraybackslash}X
                            >{\raggedright\arraybackslash}X}
\toprule
Config & Distribution-first experts & Final joint GMM & PMHHNet diagnostics \\
\midrule
Input channels   & topology, LoS, NLoS, ground mask (4 + scalar height) & topology + frozen priors + masks (9) & topology, LoS, distance, formula prior, knife-edge (5 to 6) \\
Optimizer        & AdamW ($\beta_1{=}0.9, \beta_2{=}0.999$) & AdamW & AdamW \\
Learning rate    & \num{3e-4} & \num{5e-5} & \num{1e-3} to \num{3e-4} (cosine or plateau) \\
Weight decay     & 0.01 & 0.01 & 0.015 to 0.030 \\
Batch size       & 1 & 1 & 1 \\
Grad accumulation & 8 & 8 & 16 (68) / 8 (69) \\
Warmup epochs    & 2 & 1 & 2 to 5 \\
Epochs           & 120 & 40 fine-tune epochs & 80 \\
Scheduler        & ReduceLROnPlateau, patience 15 & cosine \texttt{LambdaLR} after warmup & ReduceLROnPlateau (post-April~2025 patch) \\
SWA              & Off & Off & On, start fraction 0.60 \\
Label noise      & Off & Off & On, $\sigma_{\mathrm{dB}}=0.5$ \\
CutMix           & Off & Off (D4 augmentation only) & On, $p=0.25$ to $0.45$, FiLM-safe \\
EMA weights      & Off & Off & On, decay 0.9975 \\
Dropout          & 0.10 (decoder only) & 0.10 (decoder only) & 0.12 (69) to 0.20 (67) \\
Height encoding  & Sinusoidal FiLM, 16 log-freqs over $[12,478]$ m & Sinusoidal FiLM at encoder and decoder scales & Sinusoidal FiLM, 8 injection points \\
GMM components $K$ & 5 (Stage A, path/delay), spike+exp (angular) & 3 (residual head on anchored output) & not used (MSE/Huber heads) \\
\bottomrule
\end{tabularx}
\end{table}

Two patterns deserve emphasis. First, the move from PMHHNet diagnostics to the
distribution-first and final prior-anchored stages is a simplification, not a
complication: SWA, EMA, CutMix, label noise and the older auxiliary
stabilisation tricks are removed. The distribution-first experts use AdamW with
ReduceLROnPlateau, whereas the final joint GMM uses AdamW with the
warmup-plus-cosine \texttt{LambdaLR} schedule implemented in
\texttt{train\_try80.py}. This is consistent with the distribution-first
philosophy: once the output head and target distribution are correct, the
classical training-stabilisation tricks become unnecessary.
Second, the final model uses a smaller learning rate than the distribution-first experts
(\num{5e-5} vs \num{3e-4}) because it is fine-tuning a residual head whose
inputs already include the strong frozen priors; pushing the learning
rate higher pulls the residual head away from the anchors faster than
\texttt{prior\_guard} can catch it.

\section{Final residual GMM-head model training progression}
\label{sec:try80_training_progression}

The final residual GMM-head model was not obtained in a single run. Training began with
the joint prior-anchored residual design, but the first complete runs produced
only small improvements over the frozen priors. As a diagnostic,
this showed that the priors were already strong, but also that the residual
model was initially too constrained to correct the remaining errors, especially
for path loss.

The training path then followed four stages:

\begin{enumerate}
\item \textbf{Wide residual model.} The first full version used a shared multi-task
U-Net with \(96\) base channels and a \(128\)-dimensional FiLM condition
vector. Its residual caps were conservative, especially for path loss
(\(\pm2\) dB in LoS and \(\pm4\) dB in NLoS), and the objective retained the
complete probabilistic bundle: GMM likelihood, distribution matching,
prior anchoring, prior guard, RMSE, and MAE. This stage mainly tested whether
a residual network could improve the frozen priors without degrading them. The
multi-task sharing was deliberately placed in the encoder-decoder representation;
the PL, DS and AS residual heads remained task-specific.

\item \textbf{Wide residual model with relaxed residual bounds.} The second stage kept the same
network capacity but loosened the allowed correction range. In the selected
fine-tuned configuration, the path-loss caps are \(\pm7\) dB in LoS and
\(\pm25\) dB in NLoS, the delay-spread caps are \(\pm20\) ns in both regimes,
and the angular-spread caps are \(\pm17^\circ\) in LoS and \(\pm15^\circ\) in
NLoS. This made the residual head less constrained and produced a
clearer, but still limited, improvement.

\item \textbf{Large-capacity residual model.} The third stage increased model
capacity to \(128\) base channels and a \(160\)-dimensional conditioning
vector, while keeping the relaxed residual philosophy. It also placed more
weight on the deterministic RMSE term. This stage tested whether the remaining
gap was capacity-limited rather than only cap-limited. The larger model helped
the spread tasks, but path-loss improvement remained too small while the full
probabilistic objective was still active.

\item \textbf{Final residual fine-tuning stage.} The final stage kept the
large-capacity architecture and relaxed caps, lowered the learning rate, and
changed the objective into a short RMSE-dominant fine-tuning stage. The
probabilistic terms were disabled, the prior guard and anchor were kept only
as light stabilizers, and path loss received the strongest task weight. This
selected final model is used for the reported final results: the GMM structure
remains the residual output head, but model selection is driven by
deterministic validation performance.
\end{enumerate}

For this reason, the methodology does not treat training curves themselves as
evidence for the final claim. The relevant evidence is the same city-holdout
validation and test evaluation used throughout the thesis, reported in
Chapter~\ref{chap:results}. Training curves were used only as diagnostics for
deciding which stage should be continued or replaced.



\section{Final priors: overview and literature basis}
\label{sec:priors_overview}

\input{prior_detail_overview}

\section{Path Loss Prior}
\label{sec:try78_detail}

\subsection*{Reader map}

This section is intentionally detailed because it is the mathematical
core of the final path-loss result. A reader only interested in the modelling
logic can follow it in four passes. First, the LoS branch starts from FSPL and
adds an enhanced two-ray field sum, which explains the radial ring structure
observed in the CKM maps. Second, the NLoS branch uses a calibrated
regime-wise model over topology and obstruction features, because a LoS-shaped
formula cannot explain blocked streets by itself. Third, both branches are
fitted only on the training split and then evaluated with the same
pixel-weighted ground-mask RMSE used elsewhere in the thesis. Finally, the
prior output becomes the frozen path-loss anchor consumed by the final residual GMM-head model.

The most important results to retain while reading the details are the LoS
waterfall in Table~\ref{tab:try78_los_waterfall} and the final path-loss
numbers in Table~\ref{tab:try78_results}: the enhanced two-ray branch reduces
LoS error to about \SI{1.75}{dB}, and the full hybrid prior reaches about
\SI{1.92}{dB} overall on the held-out validation+test cities.

\input{prior_detail_try78}

\section{Delay and Angular Spread Priors}
\label{sec:try79_detail}

\input{prior_detail_try79}

\section{Joint Prior-Anchored Residual GMM-Head Model}
\label{sec:try80_detail}

\input{prior_detail_try80}
