\subsection{Physical motivation for log-domain modeling}

Delay spread (DS) and angular spread (AS) are non-negative, often
zero-inflated, and heavy-tailed. Their empirical distributions are commonly
modeled in the logarithmic domain in standard channel models
\cite{tr38901,winner2}. For a spread variable $y$, define:
\begin{equation}
z(x) = \log(1 + y(x)),\qquad \hat y(x) = \exp(\hat z(x)) - 1
\label{eq:log_transform_79}
\end{equation}

The addition of 1 is not cosmetic. It keeps exact zeros finite, avoids
\(\log(0)\), and makes the same transform usable for both delay spread in
nanoseconds and angular spread in degrees. If two pixels have native spread
values \(y_a\) and \(y_b\), their difference in transformed space is
\begin{equation}
z_a-z_b
=
\log\!\left(\frac{1+y_a}{1+y_b}\right).
\label{eq:log_ratio_79}
\end{equation}
Thus, a fixed error in \(z\) corresponds approximately to a fixed
multiplicative error in \(1+y\). This is more appropriate than native-space
least squares for spreads because large NLoS tails would otherwise dominate
the fit and the near-zero LoS mass would be ignored. The calibrated spread
prior is still a deterministic log-domain ridge estimator; it is not claiming
to estimate the full spike-plus-tail probability distribution.

\subsection{Pixel-wise notation for the spread-prior fit}

Let \(n\) index a valid ground pixel from a training map. For a selected
metric \(m \in \{\mathrm{DS},\mathrm{AS}\}\), define the native target
\(y_{m,n}\), the transformed target \(z_{m,n}=\log(1+y_{m,n})\), and the
raw log-prior value \(z^{(0)}_{m,n}\). The calibrated prior is linear in a
23-dimensional feature vector:
\begin{equation}
\hat z_{m,n}
=
\textbf{x}_{m,n}^{\top}\textbf{w}_{m,r(n)},
\qquad
\textbf{x}_{m,n}\in\mathbb{R}^{23},
\qquad
\textbf{w}_{m,r}\in\mathbb{R}^{23}.
\label{eq:try79_pixel_linear}
\end{equation}
The regime \(r(n)\) is a discrete key containing the metric, topology class,
LoS/NLoS state and UAV-height bin. In other words, the spread prior does not fit one
global line through all pixels. It fits a family of small linear models, each
one specialized to a propagation regime.

\subsection{Raw spread prior in log domain}

\begin{equation}
\begin{aligned}
z^{(0)}(x)
={}&
\log\!\left(1+b^{\mathrm{native}}_{\mathrm{LoS/NLoS}}(x)\right)
+ b_{\mathrm{topology}}
+ a_1\log(1+d_{2D})
+ a_2\theta_{\mathrm{inv}}
\\
&+
a_3\rho_{41}
+ a_4 h_{41}
+ a_5 n_{41}
+ a_6 n_{41}\theta_{\mathrm{inv}} .
\end{aligned}
\label{eq:raw_spread}
\end{equation}

where $\theta_{\mathrm{inv}}(x) = 1 - \theta_{\mathrm{norm}}(x)$ (high value at low elevation). Example coefficients for delay spread: the native anchors are $b^{\mathrm{native}}_{\mathrm{LoS}} = 4.0$ ns and $b^{\mathrm{native}}_{\mathrm{NLoS}} = 11.0$ ns, which the implementation converts to $\log(1+4.0)$ and $\log(1+11.0)$ before adding the remaining log-domain terms. The delay coefficients are $a_1 = 0.11$, $a_2 = 0.62$, $a_3 = 0.48$, $a_4 = 0.32$, $a_5 = 0.86$, $a_6 = 0.60$. After Eq.~\eqref{eq:raw_spread}, the raw log prior is clipped to $[0,8]$, inverse transformed with $\exp(\cdot)-1$, and clipped to the metric upper bound. For reproducibility, the \(z^{(0)}(x)\) entries used later in \(\textbf{x}_{79}\) denote the log value recovered from this clipped native raw prior, i.e.,
\(\log(1+\mathrm{raw\_prior}_{\mathrm{clipped}}(x))\), not an unconstrained
copy of Eq.~\eqref{eq:raw_spread}.

The raw prior is not a 3GPP formula copied directly into CKM. Instead, it
takes the 3GPP/WINNER lesson that spread statistics should be separated by
propagation state and represented in log space \cite{tr38901,winner2}, then
adds CKM-specific map morphology terms. UAV measurement studies also motivate
the elevation-angle dependence: low-elevation links usually interact with more
rooftops and street canyons than near-overhead links \cite{cai2019,izydorczyk2019}.
For reproducibility, Table~\ref{tab:try79_raw_constants} lists the fixed
raw-prior constants used before the regime-wise ridge calibration.

\begin{table}[H]
\centering
\caption{Fixed constants used in the raw spread prior of Eq.~\eqref{eq:raw_spread}.}
\label{tab:try79_raw_constants}
\scriptsize
\setlength{\tabcolsep}{3pt}
\begin{tabularx}{\textwidth}{lccX}
\toprule
Quantity & Delay spread & Angular spread & Meaning \\
\midrule
\(b^{\mathrm{native}}_{\mathrm{LoS}}\) & \(4.0\) ns & \(3.0^\circ\) &
Native LoS anchor converted to \(\log(1+b)\). \\
\(b^{\mathrm{native}}_{\mathrm{NLoS}}\) & \(11.0\) ns & \(7.0^\circ\) &
Native NLoS anchor converted to \(\log(1+b)\). \\
\(a_1,\ldots,a_6\) &
\(0.11,0.62,0.48,0.32,0.86,0.60\) &
\(0.05,0.52,0.60,0.38,0.58,0.35\) &
Coefficients of \(\log(1+d_{2D})\), \(\theta_{\mathrm{inv}}\),
\(\rho_{41}\), \(h_{41}\), \(n_{41}\), and
\(n_{41}\theta_{\mathrm{inv}}\), respectively. \\
\bottomrule
\end{tabularx}
\end{table}

The topology offset \(b_{\mathrm{topology}}\) is also fixed before ridge
calibration. For delay spread, the offsets for
\texttt{open\_sparse\_lowrise}, \texttt{open\_sparse\_vertical},
\texttt{mixed\_compact\_lowrise}, \texttt{mixed\_compact\_midrise},
\texttt{dense\_block\_midrise}, and \texttt{dense\_block\_highrise} are
\(0.00, 0.18, 0.06, 0.22, 0.24, 0.34\). For angular spread, the corresponding
offsets are \(0.00, 0.09, 0.05, 0.16, 0.18, 0.24\). These are log-domain
offsets, not native ns/degree offsets.

Every term in Eq.~\eqref{eq:raw_spread} has a physical interpretation:
\begin{itemize}
\item \(b_{\mathrm{LoS/NLoS}}\) is the base spread level for the propagation
      state. NLoS starts higher because blocked paths arrive after additional
      reflections and diffractions.
\item \(b_{\mathrm{topology}}\) is a morphology offset. Dense and high-rise
      topologies receive larger offsets because they contain more reflectors.
\item \(\log(1+d_{2D})\) gives a smooth range dependence. Longer links have
      more opportunity to encounter delayed or angularly separated paths.
\item \(\theta_{\mathrm{inv}}\) increases at low elevation angle, where the
      UAV-to-ground ray intersects more urban structure before reaching the
      receiver.
\item \(\rho_{41}\), \(h_{41}\), and \(n_{41}\) describe local building
      density, mean building height, and NLoS support in a \(41\times41\)
      window. These are direct map-derived proxies for multipath richness.
\item \(n_{41}\theta_{\mathrm{inv}}\) is an interaction term. It activates
      when low-elevation geometry and local obstruction occur together.
\end{itemize}

\subsection{Calibrated spread prior}

\subsubsection{Stage 2: 23-feature vector and design matrix}

\begin{equation}
\textbf{x}_{79}(x) = \begin{bmatrix}
z^{(0)}(x)^2,\ z^{(0)}(x),\ \log(1+d_{2D}),\ \theta_{\mathrm{norm}},\ \theta_{\mathrm{inv}}\\
h_{\mathrm{norm}},\ h_{\mathrm{norm}}^2,\ \rho_{15},\ \rho_{41},\ h_{15},\ h_{41}\\
n_{15},\ n_{41},\ n_{41}\log(1+d_{2D}),\ \rho_{41}\theta_{\mathrm{inv}},\ b_{41},\ 1\\
c_{41},\ t_{41},\ \theta_{\mathrm{norm}}\rho_{41},\ c_{41}\theta_{\mathrm{inv}},\ t_{41}\rho_{41},\ t_{41}n_{41}
\end{bmatrix}^\top
\label{eq:x79}
\end{equation}

The vector in Eq.~\eqref{eq:x79} can be grouped into five blocks:
\begin{equation}
\textbf{x}_{79}(x)
=
\left[
\underbrace{\textbf{x}_{\mathrm{prior}}(x)}_{\text{raw prior}},
\underbrace{\textbf{x}_{\mathrm{geom}}(x)}_{\text{range/elevation/height}},
\underbrace{\textbf{x}_{\mathrm{morph}}(x)}_{\text{local buildings}},
\underbrace{1}_{\text{bias}},
\underbrace{\textbf{x}_{\mathrm{int}}(x)}_{\text{interactions}}
\right]^{\top}.
\label{eq:x79_groups}
\end{equation}
The first block lets the model bend the raw prior quadratically. The geometry
block handles range and height. The morphology block measures whether there
are nearby scatterers. The interaction block allows the model to express
statements such as ``density matters more at low elevation'' or
``density-weighted transmitter clearance matters differently at low elevation''.
As in the path-loss-prior vector, some of these terms are not independent.
For instance, \(z^{(0)}(x)^2\) is determined by \(z^{(0)}(x)\), and
\(t_{41}(x)n_{41}(x)\) is tied to both the taller-building and NLoS-support
features. This cross-correlation means individual fitted coefficients should
not be interpreted as isolated physical constants. The terms were kept because
the calibrated prior improved with them, and a rigorous feature-pruning study
would require many repeated calibration runs. A simple pairwise screening would
already scale as \(O(n^2)\) for an \(n\)-term vector, before considering
higher-order subsets.

\begin{table}[H]
\centering
\small
\begin{tabularx}{\textwidth}{p{3.2cm} X X}
\toprule
\textbf{Term} & \textbf{Mathematical role} & \textbf{Physical interpretation} \\
\midrule
\(z^{(0)}(x)^2\) & Quadratic raw-prior basis. & Allows the calibration to correct curvature in the first spread estimate. \\
\(z^{(0)}(x)\) & Linear raw-prior basis. & Keeps the calibrated model tied to the log-domain physical prior. \\
\(\log(1+d_{2D})\) & Smooth distance basis. & Longer links can produce wider delay and angle supports because more scatterers can contribute. \\
\(\theta_{\mathrm{norm}}\) & Normalized elevation angle. & Separates overhead links from grazing links. \\
\(\theta_{\mathrm{inv}}\) & Inverse elevation angle. & Gives a large value to low-elevation links, where A2G models report stronger obstruction sensitivity \cite{khawaja_survey,tr38901}. \\
\(h_{\mathrm{norm}},h_{\mathrm{norm}}^2\) & UAV-height basis and quadratic height basis. & Allows the model to learn a nonlinear effect of transmitter height on spread. \\
\(\rho_{15},\rho_{41}\) & Fine/coarse local building density. & Measures how many nearby reflectors and blockers exist at two scales. \\
\(h_{15},h_{41}\) & Fine/coarse density-weighted building height. & Distinguishes low-rise clutter from high-rise clutter. \\
\(n_{15},n_{41}\) & Fine/coarse local NLoS support. & Measures how much nearby ground is geometrically shielded. \\
\(b_{41}\) & Blocker-depth term. & Measures how far nearby buildings protrude above receiver height. \\
\(1\) & Bias term. & Gives each regime its own additive offset. \\
\(c_{41}\) & Transmitter-clearance term. & Density-weighted window-area mean of UAV clearance above nearby building pixels. \\
\(t_{41}\) & Taller-building area fraction. & Fraction of the full local window occupied by buildings whose height exceeds the UAV height. \\
\bottomrule
\end{tabularx}
\caption{Term-by-term inventory of the direct terms in the spread calibration vector.}
\label{tab:x79_term_inventory}
\end{table}

\begin{table}[H]
\centering
\small
\begin{tabularx}{\textwidth}{p{3.2cm} X X}
\toprule
\textbf{Term} & \textbf{Mathematical role} & \textbf{Physical interpretation} \\
\midrule
\(n_{41}\log(1+d_{2D})\) & NLoS-distance interaction. & Lets range matter differently in blocked streets than in open ground. \\
\(\rho_{41}\theta_{\mathrm{inv}}\) & Density-elevation interaction. & Activates when dense morphology and low elevation occur together. \\
\(\theta_{\mathrm{norm}}\rho_{41}\) & Overhead-density interaction. & Allows dense areas to behave differently when viewed from high elevation. \\
\(c_{41}\theta_{\mathrm{inv}}\) & Clearance-elevation interaction. & Lets density-weighted rooftop clearance matter differently at grazing elevation. \\
\(t_{41}\rho_{41}\) & Tall-building-density interaction. & Highlights dense high-rise blocks where a large window area contains buildings above the UAV. \\
\(t_{41}n_{41}\) & Tall-building-NLoS interaction. & Activates in blocked regions where taller-building area is also present. \\
\bottomrule
\end{tabularx}
\caption{Interaction terms in the spread calibration vector.}
\label{tab:x79_interaction_inventory}
\end{table}

\begin{figure}[H]
\centering
\begin{tikzpicture}[
  >=Stealth,
  box/.style={draw, thick, rounded corners=3pt, align=center,
              minimum height=0.85cm, font=\footnotesize},
  group/.style={box, minimum width=2.9cm, fill=blue!10},
  fitbox/.style={box, minimum width=3.2cm, fill=green!12},
  arr/.style={->, thick}
]
\node[group] (prior) at (0,0) {Raw-prior\\terms};
\node[group] (geom) at (3.4,0) {Range, elevation\\and height};
\node[group] (morph) at (6.8,0) {Density, height\\and blockage};
\node[group] (inter) at (10.2,0) {Interaction\\terms};
\node[fitbox] (vec) at (5.1,-1.8) {Feature vector\\$\textbf{x}_{79}(x)\in\mathbb{R}^{23}$};
\node[fitbox] (ridge) at (5.1,-3.6) {Regime ridge fit\\$\hat z=\textbf{x}_{79}^{\top}\textbf{w}_{r}$};
\node[fitbox, fill=orange!15] (native) at (5.1,-5.4) {Native prior\\$\hat y=\exp(\hat z)-1$};
\draw[arr] (prior) -- (vec);
\draw[arr] (geom) -- (vec);
\draw[arr] (morph) -- (vec);
\draw[arr] (inter) -- (vec);
\draw[arr] (vec) -- (ridge);
\draw[arr] (ridge) -- (native);
\end{tikzpicture}
\caption[Feature grouping for the calibrated spread prior.]%
{Feature grouping used by the calibrated spread prior.
The calibrated model is linear in \(\textbf{x}_{79}\), but the features include
nonlinear transforms and interactions selected from propagation intuition.}
\label{fig:try79_feature_groups}
\end{figure}

The new features beyond those in \(\textbf{x}_{78}\) are listed below. The
shared morphology notation follows the path-loss-prior construction in
Sec.~\ref{sec:nlos_branch}. In particular, \(b(\cdot)\) is the building mask,
\(\rho_k\) is the building-density box mean, \(h_k\) is the density-weighted
building-height box mean, \(n_k\) is the ground-NLoS support box mean, and
\(\theta_{\mathrm{norm}}\) is the normalized elevation angle. The same
reflected-padding \(\mathrm{BoxMean}_k\) operator is used here, and
\(H_{\mathrm{norm}}=\SI{90}{m}\) is the building-height normalizer. This keeps
the spread-prior vector self-contained while preserving the common feature
definitions used by the path-loss prior.

\paragraph{Normalised transmitter height.}
\begin{equation}
h_{\mathrm{norm}} =
\min\!\left\{1,\,
\frac{\log\!\left(1+\max(h_{\mathrm{tx}}-h_{\mathrm{rx}},0)\right)}{\log(401)}
\right\}
\label{eq:hnorm}
\end{equation}
This compresses UAV height on a log scale and saturates it in \([0,1]\), using
\(\log(401)\) as the scale associated with a 400 m height difference. Samples
above that relative height are assigned \(h_{\mathrm{norm}}=1\), matching the
implementation. This cap is intentional: the transmitter-height distribution has
only a small high-altitude tail, so after splitting samples into
topology-specific experts there is not enough support above 400 m to estimate a
separate high-altitude response reliably. The vector also includes
\(h_{\mathrm{norm}}^2\), which is an
intentionally correlated basis expansion that lets the linear model learn a
curved height effect.

\paragraph{Blocker depth.}
\begin{equation}
b_{41}(x) = \frac{\mathrm{BoxMean}_{41}\!\left(\max(\mathrm{topology}(\cdot)-h_{\mathrm{rx}},0)\,b(\cdot)\right)(x)}{H_{\mathrm{norm}}}
\label{eq:b41}
\end{equation}
Measures how much buildings in the $41\times41$ window protrude above the receiver height.
Only building pixels contribute because of the multiplier \(b(\cdot)\). A tall
building directly next to the receiver therefore increases \(b_{41}\), while
flat open ground leaves it at zero.

\paragraph{Transmitter clearance.}
\begin{equation}
c_{41}(x) = \frac{\mathrm{BoxMean}_{41}\!\left(\max(h_{\mathrm{tx}}-\mathrm{topology}(\cdot),0)\,b(\cdot)\right)(x)}{H_{\mathrm{norm}}}
\label{eq:c41}
\end{equation}
This is a density-weighted clearance feature: the clearance above each building
pixel is averaged by \(\mathrm{BoxMean}_{41}\) over the full \(41\times41\)
window and then normalized by \(H_{\mathrm{norm}}\). Ground pixels contribute
zero through the multiplier \(b(\cdot)\). A large \(c_{41}\) therefore means
both that nearby buildings are present and that the transmitter is well above
them. A small value can mean either little local building coverage, or buildings
that reach close to or above the UAV height.

\paragraph{Taller-building fraction.}
\begin{equation}
t_{41}(x) = \mathrm{BoxMean}_{41}\!\left(\mathbf{1}[\mathrm{topology}(\cdot)>h_{\mathrm{tx}}]\cdot b(\cdot)\right)(x)
\label{eq:t41}
\end{equation}
This is the fraction of the full \(41\times41\) window area occupied by
building pixels taller than the UAV, not the fraction of buildings after
conditioning on there being a building pixel. It is zero if the UAV is above
all local buildings or if the local window contains no buildings, and grows
when the link is likely to be embedded in a high-rise canyon. It is especially
useful for distinguishing dense but low-rise zones from dense high-rise zones.

\paragraph{Six-class topology definition.}
The spread prior uses the six topology labels produced by the final Try~80
prior code. For a map, let \(\bar\rho_{\mathrm{map}}\) be the fraction of
non-ground pixels and let \(\bar h_{\mathrm{map}}\) be the mean height over
non-ground pixels only. The class assignment is:
\begin{equation}
\mathrm{class}_{79}
=
\begin{cases}
\texttt{open\_sparse\_lowrise},
& \bar\rho_{\mathrm{map}}\le0.12,\ \bar h_{\mathrm{map}}\le\SI{12}{m},\\
\texttt{open\_sparse\_vertical},
& \bar\rho_{\mathrm{map}}\le0.12,\ \bar h_{\mathrm{map}}>\SI{12}{m},\\
\texttt{dense\_block\_midrise},
& \bar\rho_{\mathrm{map}}\ge0.28,\ \bar h_{\mathrm{map}}\le\SI{28}{m},\\
\texttt{dense\_block\_highrise},
& \bar\rho_{\mathrm{map}}\ge0.28,\ \bar h_{\mathrm{map}}>\SI{28}{m},\\
\texttt{mixed\_compact\_lowrise},
& 0.12<\bar\rho_{\mathrm{map}}<0.28,\ \bar h_{\mathrm{map}}\le\SI{12}{m},\\
\texttt{mixed\_compact\_midrise},
& 0.12<\bar\rho_{\mathrm{map}}<0.28,\ \bar h_{\mathrm{map}}>\SI{12}{m}.
\end{cases}
\label{eq:topo79}
\end{equation}
These labels are map-level routing keys for the spread-prior ridge models;
they are not additional per-pixel morphology features.

\subsubsection{Stage 3: regime-wise ridge regression}

The regime $r$ for the spread prior is keyed by: metric (DS or AS), topology class (6 classes), propagation state (LoS or NLoS), antenna-height bin, giving up to $2 \times 6 \times 2 \times 3 = 72$ regimes.

Valid pixels are subsampled at rate $p = 0.02$ (2\%). The ridge regression:
\begin{equation}
\hat{\textbf{w}}^{79}_{r} = \left(\textbf{X}_r^\top \textbf{X}_r + \lambda \textbf{D}^\top\textbf{D}\right)^{-1}\textbf{X}_r^\top \textbf{z}_r,\quad \lambda = 10^{-3}
\label{eq:ridge}
\end{equation}

Here \(\textbf{X}_r\in\mathbb{R}^{N_r\times23}\) stacks one row
\(\textbf{x}_{79,n}^{\top}\) per sampled pixel in regime \(r\), and
\(\textbf{z}_r\in\mathbb{R}^{N_r}\) stacks the corresponding transformed
targets. The coefficient vector
\(\hat{\textbf{w}}^{79}_{r}\) minimizes the regularized least-squares problem
\begin{equation}
\hat{\textbf{w}}^{79}_{r}
=
\arg\min_{\textbf{w}\in\mathbb{R}^{23}}
\left\|
\textbf{X}_r\textbf{w}-\textbf{z}_r
\right\|_2^2
+
\lambda
\left\|
\textbf{D}\textbf{w}
\right\|_2^2 ,
\label{eq:try79_ridge_objective}
\end{equation}
where \(\textbf{D}=\mathrm{diag}(d_1,\ldots,d_{23})\), with \(d_{23}=0\)
and \(d_i=1\) for \(i\ne23\). Thus, the final interaction term
\(t_{41}n_{41}\), which is the last entry of Eq.~\eqref{eq:x79}, is left
unpenalized by the ridge term in the frozen calibration used by the final
model. This is an implementation detail of the stored artifact, not a separate
physical assumption: a post-hoc check that instead left the bias coefficient
unpenalized produced no meaningful RMSE change. The regularizer is small, but
it prevents unstable coefficients when a regime contains few valid pixels or
when two morphology features are strongly correlated.

Term by term, \(\textbf{X}_r^\top\textbf{X}_r\) is the feature correlation
matrix for regime \(r\), \(\lambda\textbf{D}^\top\textbf{D}\) adds a small
diagonal stabilizer to every coefficient except the final
\(t_{41}n_{41}\) coefficient, and
\(\textbf{X}_r^\top\textbf{z}_r\) is the feature-target
correlation vector in log-spread space. This is ridge regression: it is the
ordinary least-squares fit plus a penalty that discourages very large
coefficients. The penalty is important here because many morphology features
are related; for example, \(\rho_{41}\), \(h_{41}\), \(b_{41}\) and \(t_{41}\)
all increase in dense high-rise areas.

Sufficient statistics are accumulated incrementally:
\begin{align}
\textbf{A}_r &\mathrel{+}= \textbf{X}_{r,s}^\top \textbf{X}_{r,s} \quad \forall s\\
\textbf{v}_r &\mathrel{+}= \textbf{X}_{r,s}^\top \textbf{z}_{r,s} \quad \forall s
\end{align}
so the full design matrix is never materialised in memory.
This is important in practice because the calibration artifacts are already
large even after compression into fitted coefficients and profiles. The path-loss
LoS calibration JSON reused by the final model is approximately 200\,000 lines, and the
spread calibration must be rerun separately for delay spread and angular
spread whenever the feature set changes.

\paragraph{Fallback hierarchy.} During fitting, each sampled pixel updates the
exact accumulator and the four broader fallback accumulators listed below. A
coefficient vector is written only for accumulators with at least
$4 \times 23 = 92$ sampled pixels. During inference, the implementation tries
the same keys in this order and uses the first one present:
\begin{enumerate}
\item exact regime: (metric, topology class, LoS/NLoS, antenna bin),
\item same topology class + LoS/NLoS + \emph{all} antenna heights,
\item same topology class + \emph{all} LoS states + all antenna heights,
\item global topology + LoS/NLoS + all antenna heights,
\item fully global fallback: (\texttt{metric}, \texttt{global},
      \texttt{all\_los}, \texttt{all\_ant}), where \texttt{all\_los} is the
      implementation key for all visibility states, not LoS-only pixels.
\end{enumerate}

\subsubsection{Stage 4: inverse transform and clipping}

\begin{equation}
\hat y(x) = \exp\!\left(\hat z(x)\right) - 1,\qquad \hat z(x) = \textbf{x}_{79}(x)^\top \hat{\textbf{w}}^{79}_{r}
\label{eq:inv_log}
\end{equation}

Regime-dependent clipping is applied after the inverse transform:
\begin{equation}
\hat y_{\mathrm{clip}}(x)
=
\min\!\left(
y_{\max,r},
\max\!\left(0,\exp(\hat z(x))-1\right)
\right),
\label{eq:try79_clip}
\end{equation}
with delay spread clipped to $[0, 400]$~ns and angular spread clipped to
$[0, 15^\circ]$ in LoS or $[0, 90^\circ]$ in NLoS. The asymmetric angular
limit is deliberate: LoS angular spread is dominated by the direct path and is
physically closer to a near-zero spike plus small reflections, whereas NLoS
angular spread can occupy a much wider support.

\subsection{Spread-prior results}

\begin{table}[H]
\centering
\caption{Delay spread: calibrated test RMSE (ns) by topology class.}
\label{tab:ds_topo}
\begin{tabular}{lccc}
\toprule
Topology class & Overall & LoS & NLoS \\
\midrule
open\_sparse\_lowrise   & 16.88 & 17.01 & 15.43 \\
open\_sparse\_vertical  & 40.84 & 40.44 & 44.08 \\
mixed\_compact\_lowrise & 17.27 & 16.65 & 18.91 \\
mixed\_compact\_midrise & 37.88 & 38.26 & 37.25 \\
dense\_block\_midrise   & 26.90 & 29.00 & 25.02 \\
dense\_block\_highrise  & 43.52 & 40.16 & 45.08 \\
\midrule
\textbf{Overall}        & \textbf{28.10} & \textbf{27.49} & \textbf{29.62} \\
\bottomrule
\end{tabular}
\end{table}

\begin{table}[H]
\centering
\caption{Angular spread: calibrated test RMSE ($^\circ$) by topology class.}
\label{tab:as_topo}
\begin{tabular}{lccc}
\toprule
Topology class & Overall & LoS & NLoS \\
\midrule
open\_sparse\_lowrise   & 10.01 & 10.30 & 6.33 \\
open\_sparse\_vertical  & 12.70 & 12.98 & 9.97 \\
mixed\_compact\_lowrise & 14.37 & 15.99 & 8.17 \\
mixed\_compact\_midrise & 15.26 & 17.85 & 9.57 \\
dense\_block\_midrise   & 15.53 & 21.20 & 8.17 \\
dense\_block\_highrise  & 14.64 & 21.33 & 9.78 \\
\midrule
\textbf{Overall}        & \textbf{13.74} & \textbf{15.28} & \textbf{8.66} \\
\bottomrule
\end{tabular}
\end{table}

\begin{figure}[H]
\centering
\begin{tikzpicture}[
    >=stealth,
    node distance=1.5cm and 2cm,
    block/.style={draw, thick, fill=blue!10, rectangle, rounded corners, minimum width=4cm, minimum height=1cm, align=center},
    input/.style={draw, thick, fill=gray!20, rectangle, minimum width=2.5cm, minimum height=0.8cm, align=center},
    branch/.style={draw, thick, fill=orange!10, rectangle, rounded corners, minimum width=3cm, minimum height=1cm, align=center}
]
\node[input] (inputs) {Geometry, masks \& multiscale morphology \\
$d_{2D}, \theta, h_{\mathrm{tx}}, \rho_{15/41}, h_{15/41}, n_{15/41}, b_{41}, c_{41}, t_{41}$};
\node[block, below=1.5cm of inputs] (raw_prior) {Stage 1: Raw Log Prior \\ $z^{(0)}(x)$ (Eq.~\ref{eq:raw_spread})};
\node[block, below=1.5cm of raw_prior] (features) {Stage 2: Feature Vector \\ $\textbf{x}_{79}(x) \in \mathbb{R}^{23}$};
\node[block, below=1.5cm of features] (ridge) {Stage 3: Regime-wise Ridge Regression \\ $\hat z(x) = \textbf{x}_{79}(x)^\top \hat{\textbf{w}}^{79}_{r}$};
\node[branch, below=1.5cm of ridge] (invert) {Stage 4: Inverse Transform \\ $\hat y(x) = \exp(\hat z(x)) - 1$};
\node[input, below=1.5cm of invert] (clip) {Calibrated Spread Prior \\ $\hat y(x)$ (Clipped)};
\draw[->, thick] (inputs) -- node[right, font=\scriptsize] {shared spatial maps} (raw_prior);
\draw[->, thick] (raw_prior) -- node[right, font=\scriptsize] {$\mathbb{R}^{H \times W}$} (features);
\draw[->, thick] (features) -- node[right, font=\scriptsize] {$\mathbb{R}^{23 \times H \times W}$} (ridge);
\draw[->, thick] (ridge) -- node[right, font=\scriptsize] {$\mathbb{R}^{H \times W}$} (invert);
\draw[->, thick] (invert) -- node[right, font=\scriptsize] {$\mathbb{R}^{H \times W}$} (clip);
\draw[->, thick, dashed] (inputs.east) -| ++(1.5,-1.5) |- (features.east);
\node[draw, dashed, fill=yellow!10, right=1.5cm of ridge, align=left, font=\footnotesize] {Regime $r$ keys:\\1. Metric (DS/AS)\\2. Topology Class\\3. LoS / NLoS\\4. Antenna Bin};
\end{tikzpicture}
\caption[Pipeline for the delay and angular spread priors.]%
{Pipeline for the delay and angular spread priors. Both spread modalities share this identical linear architecture, predicting expected spread in the log domain via regime-wise ridge regression over 23 spatial and geometric features.}
\label{fig:try79_pipeline}
\end{figure}
