This section provides the compact entry point for the ML-calibrated physical
path-loss and spread priors and the final joint residual GMM-head model
architecture. The detailed mathematical derivations and implementation-level
fitting notes follow in the same methodology chapter, so the design and the
audit trail remain together.

\subsection{High-level architecture}

The central design decision in the final model is \emph{not} to learn path
loss, delay spread, and angular spread from scratch. Instead:

\begin{itemize}
\item the path-loss prior provides a frozen prior for \texttt{path\_loss},
\item the spread-prior module provides frozen priors for \texttt{delay\_spread} and
      \texttt{angular\_spread},
\item the final residual GMM-head model learns bounded corrections around those priors.
\end{itemize}

The deterministic or interpretable structure is kept in the prior, and the deep
model is used only for the residual part that the prior cannot explain.
Figure~\ref{fig:pipeline} illustrates the overall pipeline.

This prior phase is not presented as a new universal closed-form propagation
law. Its value is more practical: it turns physics, morphology, visibility, and
height-dependent calibration into a strong auditable baseline. The final model
then follows a divide-and-conquer split of responsibility: calibrated priors
remove the stable component of the targets, and the neural residual model
focuses on the local errors that remain.

The word \emph{prior} is therefore used for three concrete estimators, not for
a vague physical hint:
\begin{itemize}
\item \textbf{LoS path-loss prior.} This is the clearest physical branch:
      \(\mathrm{PL}_{\mathrm{LoS}}=\mathrm{FSPL}+C_{\mathrm{2ray}}+b+r\).
      FSPL uses the 3D direct Tx--Rx distance; \(C_{\mathrm{2ray}}\) comes
      from a coherent direct-plus-ground-reflected field sum; \(b\) is a
      fitted height-dependent dB offset; and \(r\) is a smoothed radial
      residual table. The UAV is always centred horizontally, so the radial
      coordinate is unambiguous, but \(h_{\mathrm{tx}}\) varies by sample:
      \(\rho(h_{\mathrm{tx}})\), \(\phi(h_{\mathrm{tx}})\), \(b(h_{\mathrm{tx}})\)
      and \(r(d_{2D},h_{\mathrm{tx}})\) are fitted in \SI{5}{m} height bins,
      smoothed, and linearly interpolated at inference. It is not a generic
      free-space formula and it is not a learned linear regressor.
\item \textbf{NLoS path-loss prior.} This starts from a raw blocked-link
      propagation estimate \(P_0\), then calibrates it with local morphology:
      building density, building height, NLoS support, shadow/blocker proxies,
      elevation angle, and interactions. It uses antenna-height regimes
      (\texttt{low\_ant}, \texttt{mid\_ant}, \texttt{high\_ant}) because the
      obstruction geometry changes qualitatively when the centred UAV rises
      above the local skyline. It is deliberately separate from the LoS branch
      because blocked pixels do not follow the same two-ray mechanism.
\item \textbf{Delay- and angular-spread priors.} These are fitted in
      \(\log(1+y)\) space. The raw estimate captures propagation state,
      distance, elevation, and local morphology; the final ridge calibration
      adds the 23-feature vector of Sec.~\ref{sec:try79_detail}. That vector
      includes continuous normalized UAV height, transmitter clearance above
      nearby rooftops, and the local fraction of buildings taller than the
      transmitter. This is why the spread prior can represent both the
      near-zero LoS spike and the NLoS tail better than a native-space global
      average.
\end{itemize}

\begin{table}[H]
\centering
\caption{Term-by-term interpretation of the main frozen-prior quantities.}
\label{tab:prior_term_meanings}
\footnotesize
\renewcommand{\arraystretch}{1.08}
\begin{tabularx}{\textwidth}{>{\raggedright\arraybackslash}p{0.18\textwidth}
>{\raggedright\arraybackslash}p{0.20\textwidth}
>{\raggedright\arraybackslash}X}
\toprule
Term & Prior block & Interpretation \\
\midrule
\(d_{\mathrm{LoS}}\) & LoS path loss &
3D direct Tx--Rx distance. It sets the FSPL baseline and uses the fixed
\(h_{\mathrm{rx}}=\SI{1.5}{m}\) receiver height. \\
\(d_{\mathrm{ref}}\) & LoS path loss &
Image-source reflected distance. It is the length of the single
ground-reflected path used in the coherent two-ray correction. \\
\(C_{\mathrm{2ray}}\) & LoS path loss &
Direct-plus-reflected field correction in dB. Negative values mean constructive
gain over FSPL; positive values mean destructive fading. \\
\(\rho(h_{\mathrm{tx}})\) & LoS path loss &
Height-interpolated effective reflection amplitude. It is fitted from CKM
rather than assumed from one fixed ground material. \\
\(\phi(h_{\mathrm{tx}})\) & LoS path loss &
Height-interpolated effective reflection phase. It absorbs the phase offset of
the simplified one-reflection geometry. \\
\(b(h_{\mathrm{tx}})\) & LoS path loss &
Height-dependent dB bias remaining after the coherent two-ray fit. \\
\(r(d_{2D},h_{\mathrm{tx}})\) & LoS path loss &
Smoothed radial residual table. It corrects distance-dependent errors that
remain after FSPL, two-ray interference, and height bias. \\
\(P_0\) & NLoS path loss &
Raw blocked-link propagation estimate before map-morphology calibration. \\
\(\rho_k\) & NLoS PL and spreads &
Fraction of building pixels in a \(k\times k\) neighbourhood; a local density
proxy. \\
\(h_k\) & NLoS PL and spreads &
Local building-height summary in a \(k\times k\) neighbourhood; a proxy for
nearby reflectors and blockers. \\
\(n_k\) & NLoS PL and spreads &
Local NLoS-support fraction; indicates how much of the neighbourhood is
geometrically blocked. \\
\(s\) & NLoS path loss &
Shadow/blocker proxy used by the NLoS linear calibration. \\
\(\theta\), \(\theta_{\mathrm{norm}}\), \(\theta_{\mathrm{inv}}\) &
PL and spreads &
Elevation-angle terms. Low elevation links interact with more buildings, so
\(\theta_{\mathrm{inv}}\) is especially useful for spread tails. \\
\(b_{41}\), \(c_{41}\), \(t_{41}\) & Spreads &
Blocker depth, transmitter clearance, and local taller-building fraction; they
separate open overhead links from obstructed street-canyon-like links. \\
\bottomrule
\end{tabularx}
\end{table}

\begin{figure}[H]
\centering
\begin{tikzpicture}[
  box/.style={rectangle, draw, rounded corners=3pt, minimum width=2.2cm,
              minimum height=0.75cm, align=center, font=\footnotesize},
  arr/.style={-{Stealth[length=5pt]}, thick},
  node distance=0.5cm and 1.1cm
]
\node[box, fill=blue!15]  (t78) {Path-loss\\prior};
\node[box, fill=green!15, below=of t78] (t79) {Spread\\priors};
\node[box, fill=orange!15, right=1.2cm of t78, yshift=-0.62cm] (inject)
      {Prior injection\\(input channels)};
\node[box, fill=red!15, right=1.2cm of inject] (dnn)
      {DNN residual\\multi-task};
\node[box, fill=gray!20, right=1.2cm of dnn] (out)
      {PL, DS, AS\\outputs};

\draw[arr] (t78) -- (inject);
\draw[arr] (t79) -- (inject);
\draw[arr] (inject) -- (dnn);
\draw[arr] (dnn) -- (out);

\node[below=0.25cm of dnn, font=\scriptsize, text width=2.4cm, align=center]
      {$\delta \in [-\delta_{\max}, +\delta_{\max}]$};
\end{tikzpicture}
\caption[Overall prior-anchored pipeline with bounded DNN corrections.]
{Overall prior-anchored pipeline. The frozen priors are injected into the DNN, which learns a bounded correction $\delta$ around them.}
\label{fig:pipeline}
\end{figure}

\subsection{Prior design rules}
\label{sec:good_prior_design}

The main lesson from the prior phase is that a useful prior is not just a
physical-looking equation. It is a complete estimator: the formula, fitted
calibration, feature units, LoS/NLoS mask, topology features, and evaluation
mask must be identical between fitting and inference. Earlier prior experiments
used many of the same physical ingredients as the final path-loss prior, but they did
not always keep those pieces aligned. The final prior became strong because it
made the prior self-contained and auditable under the same city-holdout
contract used for the final model.

The distinction matters most for NLoS. A calibration fitted for one physical
formula cannot reliably correct another; local height and density features
must have the same units when fitted and when applied; and regime assignment
should come from the sample geometry rather than only from a broad city label.
Once those constraints were enforced, the prior could separate deterministic
large-scale structure from the genuinely local residual that the final model later
learned.

This does not mean that all earlier NLoS failures were only implementation
mismatches. The no-prior and weak-prior neural branches also failed because
single-output point regression was a poor match for blocked, multi-modal,
heavy-tailed NLoS and spread targets. That is why the thesis keeps both
lessons: the final priors fix the prior-estimator problem, while
the distribution-first predecessors motivate distribution-aware heads and residual
uncertainty.

\begin{table}[H]
\centering
\caption{Compact design rules extracted from the prior experiments.}
\label{tab:prior_design_rules}
\small
\begin{tabularx}{\textwidth}{>{\raggedright\arraybackslash}p{0.28\textwidth}
>{\raggedright\arraybackslash}X}
\toprule
Rule & Effect on the final pipeline \\
\midrule
Fit and apply the same formula &
The calibration corrects the actual runtime prior, not a neighbouring variant. \\
Keep units explicit &
Distances, heights, topology values, and normalized features have one shared
definition across fitting, validation, and test. \\
Use the metric mask during fitting &
The prior is optimized for valid receiver pixels rather than for building cells
that are excluded from the final metric. \\
Hard-separate LoS and NLoS &
The two regimes keep different physics, feature sets, and calibration
parameters when the dataset provides a reliable mask. \\
Let literature define axes, not all coefficients &
Distance, height, frequency, and elevation come from propagation theory;
effective reflection and local morphology corrections are fitted from CKM. \\
Learn bounded residuals after the prior &
The final model is allowed to correct local errors without overwriting the calibrated
large-scale structure. \\
Treat calibration artifacts as part of the model &
Large JSON and coefficient artifacts are not bookkeeping only; they store the
height-, density-, visibility-, and morphology-specific decisions that make the
prior a competitive baseline. \\
\bottomrule
\end{tabularx}
\end{table}

\subsection{Literature ingredients and thesis adaptations}

The final pipeline is a hybrid between classical propagation models,
distribution-aware regression, and dataset-specific calibration.
Table~\ref{tab:prior_literature_mapping} keeps the mapping explicit; the
detailed equations and fitting procedures are given in the following
path-loss-prior, spread-prior, and residual-GMM methodology sections.

\begin{table}[H]
\centering
\caption{Literature-to-design mapping used by the final prior-anchored model.}
\label{tab:prior_literature_mapping}
\scriptsize
\renewcommand{\arraystretch}{1.12}
\begin{tabularx}{\textwidth}{>{\raggedright\arraybackslash}p{0.18\textwidth}
>{\raggedright\arraybackslash}X
>{\raggedright\arraybackslash}X
>{\raggedright\arraybackslash}X}
\toprule
Block & Literature basis & CKM adaptation & New role in this thesis \\
\midrule
LoS path loss prior &
FSPL and coherent two-ray propagation. &
Effective complex reflection, height-bin interpolation, and radial residual
fitting on CKM LoS pixels. &
Reconstructs the radial interference pattern and supplies the frozen LoS
anchor. \\
NLoS path loss calibration &
COST-231/Hata-style attenuation, A2G elevation envelopes, and LoS/NLoS
separation. &
Map-derived density, height, NLoS-support, elevation, and shadowing features
with regime-wise calibration. &
Turns a broad empirical curve into a local, topology-aware NLoS prior. \\
Delay and angular spread priors &
Log-normal spread statistics from 3GPP/WINNER-style channel models and UAV
elevation trends. &
Log-domain ridge regression with local morphology features and a sparse-regime
fallback hierarchy. &
Provides stable spread anchors without pretending that spreads are smooth
deterministic path-loss maps. \\
Final residual model &
U-Net-style dense prediction, FiLM conditioning, probabilistic regression, and
mixture-density ideas. &
Prior channels, explicit masks, residual caps, anchor gates, and a shared
multi-task trunk. &
Learns only the bounded local correction and uncertainty left after the
calibrated priors. \\
\bottomrule
\end{tabularx}
\end{table}
