\chapter[State of the Art]{State of the Art and Technical Background}

Radio-map prediction sits at the intersection of classical wireless propagation
modelling, statistical channel analysis, and dense computer-vision regression.
This chapter surveys the relevant state of the art across four axes that jointly
motivate the design choices in the final residual GMM-head model: (i)~classical and empirical propagation
models, (ii)~deep learning for radio-map estimation, (iii)~physics-informed and
hybrid neural approaches, and (iv)~distributional and multi-target channel
modelling. A final section frames fair-comparison criteria and positions the final model
within the literature.

% Section separator
\section{Channel knowledge maps as the thesis context}
\label{sec:soa_ckm_context}
% Section separator

The term channel knowledge map (CKM) refers to a site-specific database of
location-tagged channel knowledge, such as path loss, channel gain, angles, or
delay information, used to make future radio systems environment-aware
\cite{zengx2021ckm,ckmtutorial2024}. This framing is more precise than a
generic radio map: the map is not only an image-like coverage estimate, but a
queryable prior that can reduce channel-state acquisition overhead, guide
beamforming, support planning, and provide initial channel knowledge before
online measurements are available.

Recent CKM literature also clarifies why the thesis uses a large offline
ray-traced dataset rather than sparse field samples. Xu and Zeng
\cite{ckmdata2024} analyse how CKM construction error depends on spatial sample
density and on the number of neighbouring samples used for online prediction,
showing the explicit cost/accuracy trade-off of measurement-driven CKM
construction. For a thesis whose goal is to learn dense \(513\times513\)
targets under city holdout, this motivates the simulation-first route: it
provides full supervision everywhere, while the final model is still evaluated
as a surrogate for a ray-traced CKM rather than as a replacement for field
calibration.

The data ecosystem has also expanded since the first path-loss challenge.
RadioMap3DSeer \cite{dataset2212} and the ICASSP~2023 challenge
\cite{icassp2023challenge} provide the closest open benchmark family for
path-loss image prediction, while the directive-antenna dataset of Jaensch et
al.\ \cite{jaensch2024directiverme} and CKMImageNet
\cite{ckmimagenet2025} show the broader move toward open, location-specific
datasets with richer channel and environment representations. The CKM dataset
used in this thesis sits in the same data-driven tradition, but differs in its
continuous UAV height conditioning, real-city holdout split, and simultaneous
path-loss, delay-spread, and angular-spread targets.

% Section separator
\section{Classical propagation models and channel statistics}
\label{sec:soa_classical}
% Section separator

\subsection{Free-space and two-ray propagation}

The most fundamental path-loss term is the \emph{free-space path loss} (FSPL),
which depends only on carrier frequency~$f$ and Tx-Rx distance~$d$:
\begin{equation}
  \mathrm{PL}_{\mathrm{FSPL}} = 20\log_{10}(d) + 20\log_{10}(f)
    + 20\log_{10}\!\left(\frac{4\pi}{c}\right).
\end{equation}
FSPL is the irreducible floor in unobstructed conditions.  In aerial scenarios a
specular ground reflection produces constructive/destructive interference
periodically with distance.  The \emph{coherent two-ray} model captures this by
adding the direct and reflected fields before converting to dB
\cite{rappaport,wocc2021}:
\begin{equation}
  E_{\mathrm{total}} = E_{\mathrm{direct}}
    + \Gamma \cdot E_{\mathrm{reflected}} \cdot e^{j\Delta\phi},
\end{equation}
where $\Gamma$ is the reflection coefficient and $\Delta\phi$ is the phase
difference determined by the path-length difference.  A calibrated two-ray model
fitted to the CKM dataset achieves approximately
$1.75\,\mathrm{dB}$ RMSE on line-of-sight (LoS) pixels; the deterministic
radial structure explains most LoS variance without any neural network.

\subsection{Empirical and semi-empirical models}

Empirical models such as the Okumura-Hata and COST-231/Hata formulas extend
FSPL by incorporating frequency, antenna heights, and city-type correction terms
\cite{rappaport,cost231}.  While computationally cheap, they are calibrated on specific
frequency bands and morphology types, limiting accuracy when the geometry
departs from the fitting dataset.

The 3GPP Technical Report 38.901 \cite{tr38901} formalises stochastic channel
models for frequencies from 500~MHz to 100~GHz for urban macro (UMa), urban
micro (UMi), rural macro (RMa), and indoor scenarios.  It provides separate
path-loss polynomials for LoS and non-LoS (NLoS) conditions, breakpoint
distances, and shadow-fading standard deviations, which serve as reference
baselines throughout this thesis.

\subsection{UAV and air-to-ground channel models}

Classical terrestrial channel models assume the transmitter is at rooftop height
or below.  Unmanned aerial vehicle (UAV) base-station scenarios introduce a
third dimension: the UAV can fly from near-ground level to several hundred
metres.  Al-Hourani et al.~\cite{alhourani2014} showed that the LoS probability
increases monotonically with elevation angle and depends on city morphology type
(suburban, urban, dense urban, high-rise).  The ITU and 3GPP A2G channel model
family \cite{tr38901,khawaja_survey} parameterises this dependence through
logistic functions of the elevation angle
$\theta = \arctan\!\bigl((h_{\mathrm{tx}} - h_{\mathrm{rx}})/d_{2D}\bigr)$,
where $h_{\mathrm{tx}}$ and $h_{\mathrm{rx}}$ are transmit and receive heights
respectively and $d_{2D}$ is the horizontal distance.

Recent work by Saboor and Vinogradov \cite{saboor2025height} makes this
height dependence especially relevant for the present thesis: in simulated
26~GHz urban A2G links, LoS path loss remains close to a free-space exponent,
whereas NLoS path-loss exponents and shadow fading evolve with ABS height and
with urban layout. Vinogradov et al.\ \cite{vinogradov2026shadow} also propose
3D shadow projections as a fast way to generate spatially consistent A2G LoS
maps without full point-by-point ray tracing. These results support two design
choices used throughout this work: treating the LoS/NLoS mask as a physical
input rather than as a cosmetic label, and conditioning every dense prediction
on UAV height.

A key finding from the CKM dataset is that LoS path-loss residuals after
two-ray subtraction form almost perfect radial rings around the transmitter.
This confirms that buildings act primarily as a \emph{visibility mask} in LoS
conditions rather than as the main source of residual path-loss variance,
validating the prior-subtraction philosophy of the final path-loss prior.

\subsection{Delay spread and angular spread in standard channel models}

Beyond path loss, full channel characterisation requires second-order statistics.
The \emph{root-mean-square delay spread} $\tau_{\mathrm{rms}}$ measures the
time-domain dispersion of multipath arrivals and drives inter-symbol interference
in wideband systems.  The \emph{azimuth angular spread} $\sigma_{\mathrm{AS}}$
at the receiver quantifies the spatial diversity of arriving paths.

3GPP~TR~38.901 \cite{tr38901} models both quantities as log-normal random
variables conditioned on the LoS/NLoS state and distance.  The WINNER~II and
COST~2100 models adopt the same convention.  In practice, the distributions of
delay spread and angular spread are right-skewed with a heavy exponential tail
and often include a spike near zero representing the dominant LoS path. The
calibrated spread-prior analysis in this thesis empirically confirms this
\emph{spike-plus-exponential} shape for
all topology classes in the CKM dataset, with skewness exceeding~5 and kurtosis
exceeding~40 in several morphology classes.  A Gaussian model cannot represent
this shape faithfully, motivating the mixture and distributional approach used
by the final residual GMM-head model.

% Section separator
\section{Deep learning for radio-map estimation}
\label{sec:soa_dl}
% Section separator

\subsection{Image-to-image translation baselines}

The radio-map prediction problem can be cast as an image-to-image regression:
given an input topology map (and optionally a transmitter location channel), the
model predicts a path-loss map of the same spatial extent.  The pix2pix framework
\cite{isola2017pix2pix} established a general recipe for such tasks using a
conditional GAN with a U-Net generator and a PatchGAN discriminator.  Early
iterations of this thesis used this formulation.  In practice the adversarial
objective sharpens spatial texture but does not address the height-dependent
radial pattern or the heavy NLoS tails that dominate the error budget in the CKM
dataset.

RadioUNet \cite{radiounet2020} is the canonical convolutional path-loss-map
baseline before the PMNet challenge family. In its accurate-map settings it
reports grey-level RMSE values around \(0.0203\)--\(0.0384\); using the paper's
\SI{80}{dB} grey-level conversion gives an approximate \SI{1.6}{dB}--\SI{3.1}{dB}
path-loss scale. RadioGUNet \cite{radiogunet2025} extends this outdoor
RadioMapSeer line with group-equivariant convolutions, reporting
\SI{1.304}{dB} on the easiest DPM setting, \SI{1.936}{dB} on the IRT setting,
and \SI{1.392}{dB} with cars. These papers are useful path-loss scale
references, but they do not report dense delay-spread or angular-spread maps and
do not use the CKM city-holdout, continuous-UAV-height contract.

\subsection{PMNet and the ICASSP outdoor challenge family}

PMNet \cite{pmnet2023} was the winner of the first ICASSP 2023 Outdoor Pathloss
Radio Map Prediction Challenge \cite{icassp2023challenge}.  It uses a deep
encoder-decoder with skip connections, trained and evaluated on the
RadioMap3DSeer dataset~\cite{dataset2212}.  RadioMap3DSeer consists of
procedurally generated city maps with transmitters fixed at rooftop height
(6 to 20~m) and building interior pixels masked to zero error.

The challenge metric is RMSE in normalised image units.  The RadioMap3DSeer
image mapping clips path gain to the range
$[\mathrm{PL}_{\mathrm{thr}}, \mathrm{PL}_{\mathrm{max}}] = [-111, -75]\,\mathrm{dB}$,
yielding a $36\,\mathrm{dB}$ dynamic window.  Converting to physical dB requires
multiplying normalised RMSE by~36.  After this correction, the true physical
errors are:

\begin{table}[ht]
\centering
\caption[Corrected ICASSP 2023 outdoor RMSE.]%
{Corrected physical RMSE for ICASSP~2023 outdoor challenge methods.
The normalised scores are taken from the challenge family and converted to dB
using the \SI{36}{dB} RadioMap3DSeer dynamic range described in the dataset
paper \cite{icassp2023challenge,dataset2212}.\protect\footnotemark}
\label{tab:icassp2023corrected}
\begin{tabular}{lcc}
\toprule
\textbf{Method} & \textbf{RMSE (norm.)} & \textbf{True RMSE (dB)} \\
\midrule
PMNet \cite{pmnet2023} & 0.0383 & 1.38 \\
PPNet baseline \cite{icassp2023challenge} & 0.0507 & 1.83 \\
\bottomrule
\end{tabular}
\end{table}
\footnotetext{The RadioMap3DSeer dynamic range is defined by the path-gain
clipping window $[\mathrm{PL}_{\mathrm{thr}},\mathrm{PL}_{\mathrm{max}}] =
[-111,-75]\,\mathrm{dB}$ used to build the normalised target image, so the
physical span is $-75-(-111)=36\,\mathrm{dB}$. Multiplying the normalised RMSE
values reported in the challenge paper \cite{icassp2023challenge,dataset2212}
by this factor converts them to physical dB, which gives
$0.0383\times 36\approx 1.38$ for PMNet.}

These numbers are not directly comparable to the CKM evaluation for two
structural reasons: (i)~fixed Tx height eliminates the continuous UAV
height-conditioning challenge; (ii)~the 36~dB window is much narrower than
CKM's effective 65 to 130~dB range, intrinsically compressing RMSE values.
The table is therefore useful as a scale reference for modern image-to-image
radio-map predictors, not as evidence that the final CKM task is solved by those
architectures. In particular, the final model must predict over unseen real-city
geometries, at native \(513\times513\) resolution, with a continuously varying
UAV altitude and with metrics computed only over valid ground receiver pixels.

A separate transfer test performed in this thesis with official PMNet checkpoints
placed official PMNet checkpoints on CKM samples.  The best zero-shot result was
approximately $84\,\mathrm{dB}$ RMSE, with most models collapsing to a nearly
constant prediction.  After oracle linear calibration from test data the minimum
achievable RMSE was still approximately~$42\,\mathrm{dB}$.  This confirms that
PMNet's strong challenge score does not transfer to the structurally harder CKM
setting.

\subsection{Transformers, state-space models, and attention-based encoders}

The RMTransformer \cite{rmtransformer2025} evaluates a hybrid MaxViT/CNN
encoder--decoder on the USC/PMNet dataset rather than on the ICASSP challenge
test set.  It reports a normalised RMSE of~0.007148 versus~0.01046 for PMNet
under a random 90/10 split on \(256\times256\) maps.  The paper normalises
received power/path loss from \(-254\) to \SI{0}{dBm}, so the reported pixel
RMSE corresponds to roughly \SI{1.82}{dB} on its own scale, and the reported
channel prediction error of 0.008099 corresponds to roughly \SI{2.06}{dB}.
This supports the use of attention for long-range urban context, but it is not
inserted in Table~\ref{tab:icassp2023corrected} because it is not measured on
RadioMap3DSeer or under the ICASSP challenge contract.

Recent attention and foundation-model variants extend the effective receptive
field of radio-map predictors \cite{wicopg2025,fmrme2026}.  They are attractive
when global context is necessary, such as tracking a shadowing corridor across
an entire $513 \times 513$ map, but they usually require substantially more data
and compute than convolutional methods to reach the same local detail quality.

\subsection{Foundation models for radio maps}

The most recent shift in the literature is towards \emph{wireless channel
foundation models} pre-trained on large heterogeneous datasets before
task-specific fine-tuning.  WiCo-PG \cite{wicopg2025} uses dual VQGANs and a
transformer with frequency-guided mixture-of-experts.  It exploits RGB drone
camera images as an auxiliary modality and achieves an NMSE of~0.012 on its
constructed multi-modal U2G benchmark.  RadioLAM \cite{radiolam2025} targets
fine-grained 3D radio-map estimation from ultra-low sampling rates using a
large generative model pipeline.  FM-RME \cite{fmrme2026} frames radio map estimation as a foundation
modelling problem with self-supervised pre-training on simulated propagation
data.

These foundation approaches represent the long-term direction of the field but
require massive training corpora and compute budgets well beyond the scope of a
single CKM dataset.  They are cited here as the horizon against which this thesis
positions its own contributions.

% Section separator
\section{Physics-informed and hybrid neural approaches}
\label{sec:soa_physics}
% Section separator

\subsection{Geometry-assisted and diffraction-aware deep learning}

Purely data-driven models must implicitly learn diffraction geometry from the
topology map.  An alternative is to pre-compute a physics-based channel prior and
provide it as an additional input channel, reducing the learning problem to
residual estimation.

Shadow-projection methods \cite{vinogradov2026shadow} occupy an intermediate
point between stochastic LoS-probability models and full ray tracing. They use
3D building geometry to project blocked ground regions from the aerial
transmitter, producing a deterministic LoS map that is spatially consistent
along user trajectories. The CKM dataset already provides a ray-traced LoS
mask, so this thesis does not need to generate the mask itself; nevertheless,
the method is directly relevant because it explains why a binary visibility
channel can carry so much of the structure that the neural network otherwise
has to infer from the height map alone.

Geometry-assisted diffraction features~\cite{geomDL2024} use virtual obstacles,
multi-screen knife-edge propagation, and scattering-aware local geometry cues to
inject physically meaningful NLoS information into a deep model.  This is
closely related to the motivation behind the thesis diffraction-prior
experiments, although the implementation here is a lightweight per-map channel
rather than the full geometry model-assisted learning framework of that paper.
The final pipeline ultimately relies on the stronger calibrated NLoS prior
rather than on a standalone knife-edge ablation.

\subsection{Physics-informed neural networks}

The ReVeal architecture \cite{reveal2025} operates as a physics-informed neural
network (PINN) by incorporating a second-order PDE residual as an auxiliary loss
term.  This penalises predictions with inconsistent local spatial curvature
under the derived propagation constraint, reducing corner-transition artefacts.
ReVeal achieves~$1.95\,\mathrm{dB}$ RMSE in rural scenarios using only 30
training points by leveraging the physical constraint as a strong regulariser.
The PMHHNet diagnostic phase in this thesis implemented a lightweight analogue:
a masked Laplacian penalty on the residual prediction, which penalises locally
non-smooth outputs without solving the full ReVeal operator.

\subsection{Prior-residual decomposition}

A principled hybrid approach is to \emph{freeze} a calibrated analytic prior and
train the neural component to predict only the structured residual.  This design
reduces the effective learning problem from the full path-loss range to a
narrower residual distribution, improving data efficiency and interpretability.

The final path-loss prior implements this philosophy: an enhanced two-ray model fitted
per height bin provides a prior map
$\hat{y}_{\mathrm{PL}}(\textbf{x}) = \mathrm{TwoRay}(d, h_{\mathrm{tx}})$,
and the neural component predicts the residual
$\epsilon = y - \hat{y}_{\mathrm{PL}}$. The final spread prior extends the idea to delay spread
and angular spread by using a ridge-regression model in log space with
TX-centred geometric features and local morphology statistics. The final
residual GMM-head model unifies all three targets under a single shared network anchored
to both priors.

% Section separator
\section{Channel statistics prediction: delay spread and angular spread}
\label{sec:soa_stats}
% Section separator

\subsection{Empirical models for spread parameters}

Standard channel models characterise delay spread and angular spread as
log-normal distributions parameterised by distance and LoS/NLoS state.
TR~38.901 \cite{tr38901} provides polynomial fits for the log-normal mean and
standard deviation as functions of 2D distance and transmitter height across UMa,
UMi, and RMa scenarios.  These fits are limited to the morphology class and
frequency band of the underlying measurement campaigns.

Log-domain modelling is physically motivated: spread parameters are
multiplicative quantities in physical space, so their logarithm is more
Gaussian-like and additive-noise assumptions are better justified. The calibrated spread prior adopts
this convention by working in $\log(1+y)$ space throughout, consistent with the
3GPP empirical modelling framework.

\subsection{Deep learning for wideband channel statistics}

Work on deep learning for delay and angular spread is sparser than for path
loss. Most published methods either predict spread globally per sample (a single scalar)
rather than per pixel, or treat it as an ancillary output of a multitask
path-loss model.  Per-pixel dense prediction of delay spread and angular spread
over $513 \times 513$ maps is, to the best of the author's knowledge, not
addressed by published methods at the time of writing.

The closest external quantitative references for spreads are therefore scalar
channel-characteristic predictors rather than dense CKM-map benchmarks. Huang et al.\
\cite{huang2025a2gtransformer} are not used in the results comparison tables
because their Wireless InSite campus dataset predicts scalar A2G link
characteristics from coordinates and link features, not dense CKM maps from
city topologies.

The main challenge is the spike-plus-exponential distribution shape of spread
targets.  A model trained only with mean-squared error (MSE) collapses to the
conditional mean, which lies in the interior of the spike and the exponential
tail and is therefore a poor representative of either regime. The
distribution-first spread predecessor addressed this with a Gaussian mixture
model (GMM) distribution head, and the final residual GMM-head model incorporates the
log-domain ridge prior as an anchor before training the residual neural head.

\subsection{Distributional mismatch and the spike-plus-exponential problem}

The spike-plus-exponential shape, consisting of a Dirac-like concentration near zero
(the LoS-dominated case) plus a heavy exponential tail (NLoS multipath), was
confirmed empirically on the CKM dataset.  Skewness exceeds~5 and excess kurtosis
exceeds~40 for both delay and angular spread in several topology classes.  A
single Gaussian or even a symmetric loss function does not fit this family.

Mixture models are the natural solution, but the required number of components
depends on what is being modelled. The histogram study selected the
best family for the raw target distributions, where path loss and the spread
targets often needed four or five components to match the full marginal shape.
The final model uses three components for a different object: the residual distribution
after the frozen priors have already explained the dominant LoS,
NLoS, delay-spread, and angular-spread structure.  This is a small design
change rather than a contradiction: the prior removes much of the target-side
multi-modality, so the residual head only needs a compact mixture to represent
under-correction, near-zero correction, and over-correction modes.  This
motivates the Stage-A distribution head in the path-loss and spread
predecessors, and the lighter GMM-3 residual head carried into the final model.

% Section separator
\section{Distribution-aware and multi-target modelling}
\label{sec:soa_distribution}
% Section separator

\subsection{Heteroscedastic uncertainty}

Kendall and Gal \cite{kendall2017uncertainties} established the framework for
heteroscedastic neural networks in computer vision: predict both a mean output
$\mu(x)$ and a log-variance $\log \sigma^2(x)$, optimising the negative
log-likelihood (NLL)
\begin{equation}
  \mathcal{L} = \frac{(\mu - y)^2}{\sigma^2} + \log \sigma^2.
\end{equation}
Pixels with high aleatoric uncertainty (deep NLoS shadows, building edges)
receive lower effective weight, while the network is penalised for
overestimating uncertainty everywhere. A PMHHNet uncertainty diagnostic applied this
loss to path-loss prediction, with $\sigma^2$ predicted as a second output
channel of PMHHNet.

\subsection{Gaussian mixture model heads}

Beyond heteroscedastic Gaussians, the GMM head allows multi-modal output
distributions.  Mixture-density networks have already appeared in wireless
channel modelling: Saleh et al.\ \cite{saleh2021probabilistic} use an MDN to
predict probabilistic mmWave path-loss distributions, Garcia Marti et al.\
\cite{garciamarti2020mixture} learn a stochastic Gaussian-mixture channel
model for physical-layer design, and Lee et al.\ \cite{lee2024timevarying}
derive conditional received-power PDFs with a mixture-density network.  These
works support the use of mixture outputs for wireless propagation, but they
operate at the level of scalar or channel-sample distributions rather than
dense radio-map generation.

The distribution-first path-loss predecessor implements a two-stage architecture: Stage~A produces per-sample GMM
parameters $\{\pi_k, \mu_k, \sigma_k\}_{k=1}^{K}$ via a shallow convolutional
encoder with sinusoidal FiLM conditioning on UAV height; Stage~B is a small
U-Net that produces per-pixel component assignments $p_k(i,j)$ and
within-component deviations $z(i,j)$.  The final prediction is
\begin{equation}
  \hat{y}(i,j) = \sum_{k=1}^{K} p_k(i,j) \cdot \bigl(\mu_k + z(i,j) \cdot \sigma_k\bigr).
\end{equation}
This formulation avoids the mode-collapse failure of MSE regression while
maintaining a compact architecture.  The same two-stage GMM design is used in
the spread predecessor for delay and angular spread, with modified parameters for the
spike-plus-exponential prior structure.  The innovative aspect is therefore
not the isolated use of a Gaussian mixture, but its role as a global
distribution head that tells the dense decoder what value distribution the map
should contain before the model solves the harder pixel-placement problem.

\subsection{Stochastic diffusion models}

Conditional diffusion models applied to radio maps \cite{radiodiff2025} generate
physically plausible stochastic variations in deep shadow zones rather than
predicting a single blurry mean.  The iterative denoising procedure can be
conditioned on topology maps, transmitter locations, and obstacle prompts.
RadioDiff reports lower normalised RMSE and higher SSIM/PSNR than CNN and
sequence-model baselines on static and dynamic RadioMapSeer settings, but its
metrics are image-normalised rather than physical dB values.  These models are
computationally expensive at inference time and are listed as a future-work
horizon for this thesis.

\subsection{Multitask and joint prediction}

Training a single shared backbone across multiple output targets can improve
sample efficiency and consistency between targets.  Path loss, delay spread, and
angular spread are physically correlated: a deep shadow corresponds to high
delay spread and potentially wide angular spread from diffracted paths.  A joint
model can therefore use the intermediate feature maps learned for path loss to
partially explain spread patterns, reducing the effective learning problem.

The final residual GMM-head model instantiates this idea explicitly: one shared encoder processes the
topology map and UAV height, and three parallel decoder heads predict path-loss
residual, delay-spread residual, and angular-spread residual respectively, all
anchored to their corresponding physical priors.  The multi-task loss is a
weighted sum of per-target NLL terms, with weights tuned on the validation set.

% Section separator
\section{Training stabilisation and regularisation techniques}
\label{sec:soa_training}
% Section separator

\subsection{Stochastic weight averaging}

Stochastic weight averaging (SWA) \cite{izmailov2018swa} averages model weights
along the training trajectory at a low, approximately constant learning rate.
This finds wider optima in the loss landscape that generalise better to
unseen data. The CKM city-holdout evaluation makes generalisation particularly
important since validation and test cities are never seen during training.
The PMHHNet diagnostic phase incorporated SWA starting from 60\,\% of the training budget.

\subsection{Exponential moving average weights}

Separate from SWA, exponential moving average (EMA) tracking of the model
weights at a decay rate of~0.9975 produces a smoother checkpoint suitable for
validation scoring.  EMA weights are used for validation and final test inference
but not for gradient updates.  The EMA-smoothed validation score is also used as
input to the ReduceLROnPlateau scheduler to avoid reacting to noisy single-epoch
validation spikes.

\subsection{Feature-wise linear modulation}

Feature-wise linear modulation (FiLM) \cite{perez2018film} conditions
the activations of a neural network on a scalar or vector input by applying
learned affine transforms $\gamma(c) \cdot \textbf{h} + \beta(c)$ to
intermediate feature maps.  Sinusoidal positional encoding of the UAV height
scalar (using log-spaced frequencies over the full 12 to 478~m range) is fed into
FiLM layers at multiple encoder depths.  This ensures that the UAV altitude
continuously modulates the activation statistics rather than being handled as a
discrete bin, which is critical when the effective propagation regime changes
substantially with height.

% Section separator
\section{Key benchmarks and fair-comparison protocol}
\label{sec:soa_benchmarks}
% Section separator

\subsection{Why headline numbers cannot be compared directly}

The most widespread mistake in radio-map literature reviews is comparing
headline RMSE values without verifying whether the underlying tasks are
equivalent.  The following evaluation axes are the minimum required for a
valid comparison:

\begin{enumerate}
  \item \textbf{Split:} city-holdout vs.\ in-distribution random split vs.\
        same-scene split.
  \item \textbf{Transmitter height:} fixed rooftop, discrete bins, or
        continuous UAV altitude.
  \item \textbf{Building pixels:} excluded from loss, excluded from metrics,
        set to zero error, or included.
  \item \textbf{Target quantity:} path loss (dB), path gain (dBm), received
        power, delay spread, angular spread.
  \item \textbf{Dynamic range and clipping:} determines the RMSE floor and
        ceiling independently of model quality.
  \item \textbf{Inference extras and calibration source:} test-time
        augmentation (TTA), SWA, EMA weights, and whether calibration is
        train-only, validation-based, or fitted from test/field measurements.
\end{enumerate}

This thesis uses a strict city-holdout split, continuous UAV height, ground-only
metrics (building pixels excluded from loss and metrics), native
$513 \times 513$ pixel maps, and no test-data or field calibration after the
train-calibrated priors and selected final model are frozen. Calibration is
therefore not shown as a binary column in Table~\ref{tab:fairness}: the relevant
question is not whether a method has fitted parameters, but whether it adapts
to the evaluation scene using validation, test, or field measurements.
Table~\ref{tab:fairness} summarises the structural axes that can be compared
compactly.

\begin{table}[ht]
\centering
\caption{Evaluation condition comparison across representative methods.}
\label{tab:fairness}
\small
\begin{tabularx}{\textwidth}{Xcccc}
\toprule
\textbf{Method} & \textbf{Unseen city} & \textbf{Cont.\ UAV height}
  & \textbf{$\ge$512\,px} & \textbf{Comparable?} \\
\midrule
PMNet / ICASSP~2023 \cite{pmnet2023} & No & No & No & No \\
RMTransformer \cite{rmtransformer2025} & No & No & No & No \\
RadioGUNet \cite{radiogunet2025} & Partial & No & No & Partial \\
AIRMap \cite{airmap2025} & Partial & No & No & Partial \\
Gao et al.\ \cite{gao2026} & No & No & Partial & Partial \\
Tarhouni et al.\ \cite{tarhouni2025} & No & No & No & Physical ref. \\
PathFinder \cite{pathfinder2025} & Partial & No & No & Partial \\
\textbf{Final model (this thesis)} & \textbf{Yes} & \textbf{Yes} & \textbf{Yes}
  & \textbf{Reference} \\
\bottomrule
\end{tabularx}
\end{table}

Two consequences follow from this checklist. First, method rankings can change
substantially when the split changes from random maps to unseen cities. A model
that has learned the texture, street orientation, and height statistics of a
known city may still fail when the same transmitter-height range is evaluated
on a new morphology. For this reason, this thesis treats city holdout as a
core part of the task definition rather than as an optional robustness test.

Second, path loss, delay spread, and angular spread should not be treated as
three copies of the same regression problem. LoS path loss often contains a
strong deterministic radial component, whereas NLoS path loss depends on
blocked-link geometry and local morphology. Delay and angular spreads are even
more distributional: many pixels are near zero, but physically important tails
remain. A fair comparison therefore has to report the target, regime, masking
policy, and split together; otherwise an apparently lower RMSE can simply mean
that an easier quantity was measured.

\subsection{Tarhouni et al.: measured suburban path-loss prediction}

Tarhouni et al.\ \cite{tarhouni2025} study machine-learning-based path-loss
prediction in a sub-\SI{6}{GHz} suburban campus. It is useful because the errors
are in physical dB and tied to a concrete environment, but it is terrestrial,
lower-frequency, and not a dense UAV CKM with delay- and angular-spread targets.
Its role here is qualitative: real-environment generalisation tends to move
back toward multi-dB errors, which supports keeping city holdout as part of the
task definition.

\subsection{AIRMap: adaptive digital-twin radio maps}

AIRMap \cite{airmap2025} is an adaptive U-Net autoencoder processing
\(200\times200\) elevation maps for ultra-low-latency digital-twin
synchronisation.  It reports sub-\SI{4}{dB} RMSE against ray-traced path-gain
maps, and a separate lightweight calibration step using 20\,\% of field
measurements reduces the median measurement-domain error to about five percent.
Its strong numbers are not directly comparable to CKM because the target is path
gain rather than absolute path loss, the input is a single elevation channel,
and the transmitter configuration is not the continuous UAV-height setting used
in this thesis. Its reported \(\sim\)\SI{4}{ms} inference time on an NVIDIA L40S
is nevertheless a strong deployment reference: the value of a CKM surrogate is
not only accuracy, but producing dense maps fast enough for planning loops.

\subsection{Gao et al.: multi-layer segmentation with corridor weighting}

Gao et al.\ \cite{gao2026} propose a ResNet-based outdoor path-loss predictor
with Tx depth maps, Rx depth maps, a distance map, and a weighting map that
emphasises the direct Tx-Rx corridor.  Using the ICASSP~2023 outdoor data,
Table~III of the paper reports their method at \SI{5.59}{dB} RMSE, compared
to \SI{8.09}{dB} for PPNet, \SI{9.56}{dB} for RPNet, and \SI{8.72}{dB} for a
twelve-layer Vision Transformer baseline; Table~II reports \SI{12.19}{dB}
RMSE on the ITU Challenge 2024 large-area subset, also ahead of the same
baselines, with roughly 60\% fewer FLOPs. This formalises an internal lesson of
the CKM work: the important geometry is not the whole map uniformly, but the
subset near the propagation corridor.

The PMHHNet diagnostic phase implemented only a weaker Tx-centred radial proxy of that
idea. The final path-loss prior then solves the LoS part more cleanly with a two-ray radial prior,
while the final residual model uses the calibrated prior as an input channel. A full Gao-style
per-Rx corridor map remains future work because generating one for every
receiver pixel at \(513\times513\) resolution is substantially more expensive
than using the current Tx-centred distance and obstruction features.

\subsection{PathFinder: multi-transmitter domain adaptation}

PathFinder \cite{pathfinder2025} tackles distribution shift via disentangled
feature encoding and mask-guided low-rank attention, reporting the lowest MSE
of~$0.1068$ and RMSE of~$0.3263$ on unseen rural datasets. On the
RadioMap3DSeer DS-RPP split, its normalised RMSE of~$0.0331$ can be
approximately converted with the \SI{36}{dB} RadioMap3DSeer target window,
giving about \SI{1.19}{dB}. The unseen-rural score would be about
\SI{11.75}{dB} under the same window, but this is only an approximate scale
reading because the paper and code report normalised PNG image units and do not
state a separate rural physical target window. The explicit
domain-generalisation focus aligns with the CKM city-holdout spirit, though
PathFinder's setup involves multi-transmitter inputs and a different
normalisation convention. Techniques from PathFinder such as
transmitter-oriented mixup augmentation are listed as medium-priority future
work. The final model is not multi-transmitter, but the same concern motivates
its city-holdout split: a CKM surrogate is useful only if it works on new urban
layouts, not only on shuffled samples from the same city.

\subsection{ICASSP 2025 indoor challenge}

The ICASSP 2025 Indoor Pathloss Challenge \cite{icassp2025indoor} presents a
structurally different problem, dense wall-penetration NLoS effects in bounded
interior spaces, but its public results are a useful warning against
over-interpreting sub-\SI{2}{dB} outdoor challenge numbers.  The official result
page reports that the winning SIP2Net submission obtains a weighted-average RMSE
of \SI{9.411}{dB}; IPP-Net reaches \SI{9.501}{dB}; TerRaIn reaches
\SI{10.325}{dB}; and TransPathNet reaches \SI{10.397}{dB}
\cite{indoor2025results,sip2net2025,ippnet2025,transPathNet2025}.  These are
not outdoor UAV results, but they show that when NLoS physics becomes harder
and less directly observable from the input map, modern neural architectures
again hit a roughly \SIrange{9}{10}{dB} error band. The comparison to the
\SI{9.41}{dB} PMHHNet reference is diagnostic, not competitive: it shows why the
final stage moved from direct regression toward strong priors, regime
separation, and residual distribution modelling.

% Section separator
\section{Emerging techniques and position of the final model}
\label{sec:soa_emerging}
\label{sec:soa_position}
% Section separator

Across the benchmarks reviewed here, the recurring $9$ to
$10\,\mathrm{dB}$ error band in harder NLoS-heavy settings reflects the
difficulty of inferring volumetric propagation from a 2D building-height map.
The most plausible routes beyond that wall are richer geometry and
diffraction priors \cite{geomDL2024}, PDE-style physics losses such as ReVeal
\cite{reveal2025}, sparse target-domain adaptation such as RadioPiT
\cite{radiopit2025}, diffusion refinement \cite{radiodiff2025}, and larger
radio-map foundation models \cite{fmrme2026}. In this thesis those options
were filtered into a narrower single-GPU design: calibrated train-only priors,
explicit LoS/NLoS and topology diagnostics, distribution-aware residual heads,
and strict city-holdout reporting.

The final model is therefore novel less because of one isolated component than
because of the combination: dense prediction of path loss, delay spread, and
angular spread on \(513\times513\) maps; frozen physics/statistical anchors;
continuous UAV-height conditioning; and a residual GMM-head that keeps the
sample-level distribution visible before decoding per-pixel maps
\cite{saleh2021probabilistic,garciamarti2020mixture,lee2024timevarying}. It
does not claim to replace ray tracing in arbitrary real deployments, to match
RadioMap3DSeer numbers under a different benchmark contract, or to act as a
frequency-general foundation model. Its claim is narrower: under the CKM
dataset contract, it reproduces unseen-city maps with physical-unit metrics
and clearer regime-level diagnostics than the earlier direct-regression family.

Table~\ref{tab:position} summarises that position against the most relevant
published work across the comparison dimensions introduced in this chapter.

\begin{table}[ht]
\centering
\caption{Positioning of the final model against the most relevant published methods.}
\label{tab:position}
\scriptsize
\renewcommand{\arraystretch}{1.12}
\begin{tabularx}{\textwidth}{>{\raggedright\arraybackslash}p{0.28\textwidth}
                              >{\raggedright\arraybackslash}p{0.24\textwidth}
                              >{\centering\arraybackslash}p{0.11\textwidth}
                              >{\raggedright\arraybackslash}X}
\toprule
\textbf{Benchmark / method} & \textbf{Reported RMSE / score} & \textbf{City holdout} & \textbf{Interpretation} \\
\midrule
RadioUNet \cite{radiounet2020} & approx.\ 1.6--3.1\,dB equivalent path-loss RMSE & No & Classic CNN radio-map scale reference; no spread targets. \\
ICASSP~2023 winner PMNet \cite{pmnet2023,icassp2023challenge} & 1.38\,dB & No & Converted from 0.0383 normalised RMSE using the RadioMap3DSeer 36\,dB target window. \\
RadioGUNet \cite{radiogunet2025} & 1.304\,dB DPM; 1.936\,dB IRT; 1.392\,dB DPM with cars & Partial & Outdoor RadioMapSeer path-loss maps with held-out test layouts, but no UAV-height or spread targets. \\
ReVeal PINN \cite{reveal2025} & 1.95\,dB outdoor rural/suburban radio-environment RMSE & No & Real outdoor sparse-measurement mapping; not dense CKM generation. \\
RMTransformer \cite{rmtransformer2025} & 0.007148 normalised pixel RMSE, approximately 1.82\,dB under its 254\,dB scaling window & No & Recent dense radio-map path-loss reference, but random 90/10 split on 256\,px USC/PMNet maps. \\
AIRMap \cite{airmap2025} & path-gain $<4$\,dB vs.\ ray tracing; field calibration reported as median percentage error & Partial & The final model is below 2\,dB PL RMSE, so it is arguably better on this broad scale, but the target is path gain rather than calibrated path loss. \\
Gao et al.\ \cite{gao2026} & 5.59\,dB on ICASSP~2023; 12.19\,dB on ITU Challenge~2024 & No & Recent outdoor PL-map reference with Tx/Rx depth and corridor weighting; the final PL RMSE is numerically lower, but the dataset, split and feature contract differ. \\
PathFinder \cite{pathfinder2025} & 0.0331 normalised RMSE on RM3D DS-RPP; 0.3263 on unseen rural & Partial & Useful outdoor/unseen-domain reference, but reported primarily in normalised image units. \\
Tarhouni et al.\ \cite{tarhouni2025} & multi-dB, sub-\SI{6}{GHz} & No & Physical outdoor reference, not dense UAV CKM. \\
\textbf{Final model (this thesis)} & \textbf{1.65\,dB on unseen test cities} & \textbf{Yes} & \textbf{Final prior-anchored residual GMM-head model; also improves delay-spread and angular-spread RMSE over the frozen priors.} \\
\bottomrule
\end{tabularx}
\end{table}
\normalsize
The final-model number in Table~\ref{tab:position} is the selected final-model
test result on the held-out city split. The prior-anchored design lets the
neural residual head converge to a narrower error floor than the
direct-regression PMHHNet family, particularly in LoS-dominated
morphology classes where the physics anchor already achieves sub-2\,dB error,
while still preserving a strict city-holdout evaluation contract.
