\chapter{Sustainability Analysis and Ethical Implications}

This chapter keeps the ETSETB sustainability structure but removes manual-style
detail. The thesis does not manufacture a physical product; its deployable result
is a software pipeline for faster CKM prediction.

\section{Sustainability summary}

\begin{table}[htbp]
\centering
\caption{Compact sustainability matrix for the thesis.}
\label{tab:sustainability_matrix}
\scriptsize
\begin{tabularx}{\textwidth}{lXXX}
\toprule
Perspective & Development of BT & Project execution & Risks and limitations \\
\midrule
Environmental & GPU training, laptop use, storage and writing. The exploratory
project exceeded roughly \SI{400}{GPU-h}, about \SI{150}{kWh} under the simple
four-GPU-node estimate. & Once trained, the generator can replace many repeated
ray-tracing queries. On the Barcelona benchmark it runs in \SI{0.88}{s} versus
\SI{99.89}{s} for the local MATLAB ray-tracing path. & Benefits depend on reuse;
full retraining for every new city, frequency or antenna configuration would
reduce the advantage. \\
Economic & Student/supervision time and shared HPC use. The original detailed
budget estimates the academic development at about 7530 EUR. & Lower marginal
planning cost if the surrogate is reused for many UAV-height or city-layout
queries. & Simulator-to-real gaps would require field measurements, calibration
and maintenance. \\
Social & Training in propagation, ML, HPC, reproducibility and scientific
reporting. & Faster planning for temporary, emergency or rural UAV coverage. &
Coverage decisions should not be based on overconfident simulated predictions or
one global RMSE. \\
\bottomrule
\end{tabularx}
\end{table}

The environmental result is therefore mixed but defensible. Training the whole
research line had a non-negligible cost, especially the failed and exploratory
models. The final model is useful only if that cost is amortized through reuse:
the same frozen priors and checkpoint can generate many dense PL/DS/AS maps at
much lower marginal cost than repeated ray tracing under the studied setup.

\section{Ethical implications and SDGs}

The main ethical obligation is to state the scope accurately. The labels come
from ray tracing, so the reported result is simulator-level generalization to
unseen cities, not certified field performance. The model is also conditioned on
one fixed CKM contract: frequency, simulator assumptions, receiver height and
antenna conventions are not varied in the final evaluation. Future operational
use would require field validation and human review, especially for emergency or
public-service deployments.

The work aligns most directly with SDG 9 (industry, innovation and
infrastructure) because it develops a faster wireless-planning tool, SDG 11
(sustainable cities and communities) because it can support temporary and
resilient connectivity planning, and SDG 13 (climate action) because a validated
surrogate can reduce repeated simulation energy. These contributions are
indirect: the thesis provides a reproducible method and generator, not an
operational deployment by itself.
