\chapter{Sustainability Analysis and Ethical Implications}

This chapter follows the ETSETB sustainability guide structure, but in a
condensed form compatible with the reduced thesis. The work does not manufacture
a physical product; its deployable result is a software pipeline for faster CKM
prediction. The analysis therefore separates the one-time academic development
cost from the possible operational use of the trained generator.

\section{Sustainability matrix}

\begin{table}[htbp]
\centering
\caption{Sustainability matrix summary for the thesis.}
\label{tab:sustainability_matrix}
\scriptsize
\begin{tabularx}{\textwidth}{lXXX}
\toprule
Perspective & Development of BT & Project execution & Risks and limitations \\
\midrule
Environmental &
GPU training, laptop use, storage, and writing. &
Reduced need for repeated ray-tracing simulations once the surrogate is trained. &
Extra retraining, larger models, or poor reuse of cached priors could increase footprint. \\
Economic &
Student and supervision time, plus shared HPC use. &
Lower marginal planning cost if the surrogate replaces repeated simulations. &
Real deployment needs field validation and maintenance. \\
Social &
Skills gained in wireless modelling, ML, HPC workflows and scientific reporting. &
Faster UAV coverage planning for temporary, rural or emergency networks. &
Bias toward represented city morphologies and risk of overclaiming simulated results. \\
\bottomrule
\end{tabularx}
\end{table}

\section{Environmental impact}

The main environmental cost during development was computation. The final
residual GMM-head model line used a four-GPU RTX~2080~Ti node for approximately
\SI{12}{h}; the distribution-first and prior-calibration phase used additional
GPU/CPU time. Across the full exploratory project, the total exceeded roughly
\SI{400}{GPU-h}. This is a one-time research overhead and includes many failed
or diagnostic experiments, not only the selected final run.

Several decisions reduced avoidable impact: reusing saved histories and frozen
calibration artifacts, evaluating city-holdout metrics from stored checkpoints,
and keeping the final priors frozen instead of recomputing them blindly inside
every neural experiment. The final design is also relatively efficient at
inference. On the Barcelona benchmark, the local MATLAB ray-tracing path took
\SI{99.89}{s} per full-resolution sample, while the complete generator path
with masks, priors, neural prediction and export took \SI{0.88}{s}. The
sustainability advantage therefore depends on reuse: if the same calibrated
generator is queried many times, the marginal cost is much lower than repeated
ray tracing; if the model is retrained from scratch for every new scenario, that
advantage weakens.

The life-cycle interpretation is therefore important. During the research phase,
the footprint is dominated by exploration: many failed or partially successful
attempts were necessary to discover that the final path should combine strong
priors with a residual GMM-head model. During the use phase, the footprint is
dominated by inference and by occasional recalibration. The project is most
sustainable when the trained model, frozen priors, cached masks and exported
checkpoints are reused across many planning queries. It would be less
sustainable if the same design were scaled by simply increasing model size or
retraining from zero for every city.

The final generator also improves reproducibility, which has an environmental
side. A documented \texttt{.npz} output, explicit mask policy, and fixed
hardware/runtime benchmark make it easier to compare changes without rerunning
expensive ray-tracing baselines. Reproducible experiments are not only a
scientific requirement; they reduce wasted computation caused by ambiguous
settings, hidden preprocessing changes or repeated debugging runs.

\section{Economic and social impact}

Economically, the main development cost is human time: literature review,
dataset handling, training/debugging, analysis, supervision and writing. The
compute cost is small compared with this time cost in an academic project, but
it is still relevant because the full exploration required many GPU-hours. In a
planning workflow, the value of the surrogate would come from reducing repeated
simulation runs and enabling faster what-if studies over UAV heights or city
layouts.

Socially, faster CKM prediction can support temporary coverage planning for
public events, rural extensions, emergency response or infrastructure
restoration. The direct users would be network planners; the indirect
beneficiaries would be users receiving more reliable wireless coverage. The
main social risk is overconfidence: a simulated city-holdout result should not
be treated as certified field performance. Errors should continue to be reported
by LoS/NLoS state, UAV-height bin and morphology group rather than hidden behind
one global RMSE.

The social value is strongest when the tool is used as decision support rather
than as an automatic decision maker. A fast surrogate can help planners compare
candidate UAV heights, identify weak-coverage zones, and decide where detailed
simulation or measurement should be spent. It should not replace domain review:
the model has been validated on simulated CKM cities and inherits the
assumptions of the ray tracer, the topology representation, and the ground-mask
definition.

There is also a fairness dimension in the morphology split. The model is tested
on unseen cities, but the training distribution is still finite and urban
morphologies are not equally represented. Dense high-rise, unusual street
canyons, informal settlements, vegetation-heavy areas or non-European building
patterns may behave differently from the dominant CKM regimes. For that reason,
future deployment should report uncertainty or at least split diagnostics by
morphology and height, instead of using the global average as the only quality
indicator.

\section{Ethical implications and SDGs}

The main ethical obligation is honesty about validity. The dataset is generated
by ray tracing, so the thesis demonstrates simulator-level generalization to
unseen cities, not real-world deployment performance. A future operational
system would require field measurements, calibration and human review,
especially before public-service or emergency use.

A second ethical issue is transparency. The final predictor is not a pure
black-box CNN: the path-loss and spread priors expose the geometric and
morphological assumptions before the residual network is applied. This makes the
system easier to audit, because a user can inspect whether an error comes from
the LoS/NLoS mask, the prior, or the learned residual correction. That audit
path is particularly relevant for emergency or infrastructure planning, where a
wrong map can lead to misplaced confidence in predicted coverage.

The work does not process personal data, but operational radio planning can
interact with sensitive contexts such as emergency response, public events or
critical infrastructure. The generator should therefore be accompanied by clear
metadata: input source, resolution, antenna height, mask-generation mode,
checkpoint version and known limitations. Keeping that metadata with the
prediction is part of the ethical responsibility of making a fast tool: speed is
useful only if the output remains traceable.

The project aligns most directly with SDG~9, because it develops a faster tool
for wireless-infrastructure planning; SDG~11, because it can support resilient
urban and temporary connectivity; and SDG~13, because a validated surrogate can
reduce repeated simulation energy. These contributions are indirect: the thesis
provides a reproducible method and generator, not an operational network
deployment by itself.

Overall, the sustainability conclusion is conditional but positive. The research
process was computation-heavy, but it produced a reusable generator that can
replace many repeated full-resolution simulations once validated for a target
domain. The strongest sustainable-use case is therefore not ``train a larger
model'', but ``reuse a documented, prior-anchored surrogate to decide where
slower simulation or real measurement is actually needed''.
