\clearpage
\chapter*{Acknowledgements}
\addcontentsline{toc}{section}{Acknowledgements}

I would like to express my sincere gratitude to my supervisors, Sergi Abadal and Evgenii Vinogradov, for their guidance throughout this project. Although I worked on this thesis with a high degree of independence, their advice was extremely valuable. I am especially grateful to Evgenii Vinogradov, for whom artificial intelligence is not his main field of research, but whose suggestions, such as encouraging me to study the two-ray model, helped me tremendously in shaping and improving this work.

I would also like to thank Enric Monte Moreno, my professor in the \textit{Machine Learning from Data} course. Although I already had some previous knowledge of machine learning and deep learning, his classes strengthened my interest in the field even further. I am also grateful for the suggestions he gave me when I discussed my TFG with him during the course.

I would also like to thank Jordi P\'erez Romero, my professor in the mobile communications part of the \textit{TSF} course. Studying mobile communications while working on this thesis helped me understand many of the concepts involved in the project much more deeply.

I would like to thank ETSETB for creating the PARS programme, which allows students to begin the Master's Degree in Telecommunications Engineering before completing their Bachelor's degree. Without this opportunity, I would not have been able to develop this TFG with the same depth of understanding or ambition. Both subjects I mentioned were of the Master's degree, and I was taking them at the same time as working on this TFG, which was a great help.

Finally, I would like to thank my family and friends for their patience and encouragement throughout the thesis.
