\chapter{Reproducibility Appendix}

\section{Detailed attempt log}

This appendix records the detailed experimental sequence used to build the
final thesis narrative. It is intentionally more operational than the main
chapters: the purpose is to show how the final design emerged from earlier
failures, not to re-argue every result.

\subsection{Early image-to-image phase}

The first implementation treated CKM prediction as an aligned
image-to-image translation problem. This was a reasonable starting point
because the inputs and outputs share the same \(513\times513\) grid. A U-Net
generator produced dense maps and a PatchGAN discriminator judged local
patches. The problem was objective mismatch. A propagation map is not a
natural image: a visually plausible texture can be numerically wrong by many
decibels. This is why later attempts reduced or removed adversarial pressure
and moved toward explicit regression losses.

Tries 9--14 introduced height conditioning and LoS/NLoS-aware evaluation.
These changes were not cosmetic. UAV height changes the path-length geometry
and therefore cannot be hidden in the sample index. Likewise, LoS and NLoS
pixels follow different propagation laws, so mixing them into one global score
made diagnosis difficult.

\subsection{Artifact removal and masked supervision}

Tries 15--22 attacked three visible problems: checkerboard artifacts, loss of
thin street and edge structure, and weak radial consistency. GroupNorm replaced
BatchNorm because the full-resolution jobs used very small batches. A distance
channel exposed the most important geometric variable directly. Try 20 isolated
the bilinear decoder change, Try 21 isolated multiscale supervision, and Try
22 combined both. In the early contract, Try 22 was the strongest clean
path-loss baseline at about \SI{16.72}{dB}.

The next methodological correction was more important than another decoder
change: pixels inside buildings are not valid receiver locations. After Try
33, the project excluded \(topology\_map \ne 0\) pixels from the loss, from
validation and test metrics, and from error-map visualization. This made the
task stricter and invalidated easy comparisons to older runs that had not used
the same mask.

\subsection{Prior-guided residual phase}

Try 41 made the residual formulation explicit:
\[
  \widehat{\mathrm{PL}} =
  \mathrm{PL}_{\mathrm{prior}} + r_\theta(x).
\]
The initial raw formula prior was poor, with validation RMSE around
\SI{67.23}{dB}. Train-only regime-aware calibration improved it to about
\SI{24.16}{dB}. This result was simultaneously encouraging and limiting:
the prior clearly contained useful information, but it was not enough to
replace learning.

Try 42 moved away from the old U-Net/cGAN family toward a PMNet-style residual
regressor. Try 49 added prior-confidence logic and robust loss weighting; its
stage-1 branch reached about \SI{18.87}{dB}, while the calibrated prior stayed
near \SI{23.57}{dB}. The main failure mode was a heavy tail: many pixels were
moderately accurate, but some NLoS pixels produced very large errors, so RMSE
stayed much larger than MAE.

Try 48 and Try 50 were useful negative evidence. Several richer NLoS priors
and structural prior variants were tested, but a candidate refined prior at
\SI{24.1777}{dB} did not beat the older \SI{23.5746}{dB} calibration, and the
best Try 50 NLoS sandbox still remained near \SI{41}{dB}. The conclusion was
that NLoS required more than a scalar correction over a LoS-shaped prior.

\subsection{Final distribution and prior phase}

Try 66 consolidated the PMHHNet family: physical channels, GroupNorm, height
FiLM, and expert routing. It reached \SI{9.41}{dB} overall, but the split
metric showed the deeper issue: \SI{3.75}{dB} LoS and \SI{31.48}{dB} NLoS.
The project was no longer failing uniformly.

Try 76 reframed NLoS as a distributional problem. Instead of predicting a
single value everywhere, the model predicted sample-level mixture structure
and then placed components spatially. This produced a much stronger NLoS path
loss result, around \SI{3.34}{dB} pixel-weighted in the DirectML summary.

Try 78 then showed that LoS path loss was even more deterministic than the
neural history suggested. The dominant residual over FSPL was a ring field,
well explained by coherent two-ray interference with fitted effective
reflection parameters. Try 79 applied a related philosophy to delay and
angular spreads: use log-domain, regime-wise calibration over geometry and
local morphology features. Try 80 is the synthesis: frozen Try 78/79 priors
plus a neural model that predicts bounded residuals.

\section{Try 80 configuration summary}

The active Try 80 configuration uses the canonical dataset
\texttt{CKM\_Dataset\_270326.h5}, image size \(513\times513\), split seed 42,
and validation/test ratios of 15\% each after city assignment. It uses
precomputed priors when available through
\texttt{try80\_precomputed\_priors.h5}, but the priors are deterministic and
can also be recomputed.

\begin{table}[ht]
\centering
\caption{Main Try 80 configuration values.}
\label{tab:try80_config_appendix}
\begin{tabularx}{\textwidth}{lX}
\toprule
Category & Value \\
\midrule
Input channels & 9: topology, LoS mask, NLoS mask, ground mask, combined
path-loss prior, LoS path-loss prior, NLoS path-loss prior, delay-spread prior,
angular-spread prior. \\
Height conditioning & Sinusoidal embedding with FiLM modulation. \\
Backbone & Four-scale GroupNorm encoder-decoder, base width 96. \\
Residual distribution & Three mixture components per task/region head. \\
Residual limits & Path loss: 2 dB LoS and 4 dB NLoS; delay spread: 30 ns LoS
and 40 ns NLoS; angular spread: 9 degrees LoS and 13 degrees NLoS. \\
Main losses & Map NLL, distribution KL, moment matching, anchor,
prior-guard, outlier budget, RMSE, and MAE. \\
Training & AdamW, learning rate \(2\times10^{-4}\), weight decay 0.01,
120 epochs, batch size 1, gradient accumulation 8. \\
\bottomrule
\end{tabularx}
\end{table}

The final prediction for each task \(t\) has the same conceptual form:
\[
  \hat y_t(x) = y_{t,\mathrm{prior}}(x) + \alpha_t(x)\Delta_t(x),
  \qquad 0 \le \alpha_t(x) \le 1.
\]
This is the central safety mechanism. The network is allowed to improve the
prior, but the residual is bounded and gated. The prior-guard term then
penalizes cases where the model moves away from the prior and becomes worse.

\section{Where the formulas are placed}

The appendix is kept as an audit trail, not as a second methodology chapter.
Detailed formulas are therefore placed where they are first used in the main
technical narrative. This avoids having two slightly different versions of the
same prior definition.

\begin{table}[ht]
\centering
\caption{Location of the main mathematical definitions.}
\label{tab:formula_location_appendix}
\begin{tabularx}{\textwidth}{lX}
\toprule
Topic & Location in the main text \\
\midrule
Ground-pixel mask and RMSE contract &
Chapter~3, Eqs.~\ref{eq:ground_mask}--\ref{eq:rmse}. \\
Residual-learning formulation &
Chapter~3, Eq.~\ref{eq:prior_residual}. \\
Height conditioning and FiLM modulation &
Chapter~3, Eqs.~\ref{eq:sin_film}--\ref{eq:film_math}. \\
Try 78 LoS path-loss prior &
Section~\ref{sec:try78_detail}, especially Eqs.~\ref{eq:fspl},
\ref{eq:field_sum}, \ref{eq:c2ray}, and \ref{eq:pl_los}. \\
Try 78 NLoS calibration &
Section~\ref{sec:nlos_branch}, especially Eqs.~\ref{eq:cost231},
\ref{eq:a2g_nlos}, \ref{eq:x78}, and \ref{eq:pl_nlos_final}. \\
Try 79 spread priors &
Section~\ref{sec:try79_detail}, especially Eqs.~\ref{eq:raw_spread},
\ref{eq:x79}, \ref{eq:ridge}, and \ref{eq:inv_log}. \\
Try 80 prior-anchored GMM &
Section~\ref{sec:try80_detail}, especially Eqs.~\ref{eq:try80_mu},
\ref{eq:try80_point}, and \ref{eq:try80_loss}. \\
\bottomrule
\end{tabularx}
\end{table}

This placement makes the appendix optional for understanding the mathematics.
A reader can follow the thesis from the main chapters alone; the appendix adds
the chronological and reproducibility context behind those choices.

\section{Final-result placeholder}

The final Try 80 checkpoint is not yet frozen. Therefore, the validation values
reported in the main chapters must be replaced by final test-split results
before submission. The thesis should keep the current wording until that
evaluation is complete: Try 80 validation results are training diagnostics, not
final scientific claims.
