\chapter{Conclusions and Future Work}

\section{Conclusions}

This thesis started as a dense deep-learning surrogate for UAV radio maps and
ended as a hybrid, prior-anchored CKM prediction pipeline. The central lesson is
that the hardest part of the problem was not drawing plausible maps, but
matching the physical and statistical structure of each target under a strict
city-holdout protocol. Architecture changes helped, but the decisive progress
came from diagnosing the target distributions, enforcing ground-only metrics,
separating LoS and NLoS behaviour, and using calibrated priors where the data
clearly followed a known propagation structure.

The original numerical targets were met by the finished ML-calibrated priors. The path-loss prior
achieved \SI{1.9246}{dB} overall path-loss RMSE on its standalone 30\%
evaluation split (\SI{1.9383}{dB} when re-evaluated as the frozen prior on the
final 14-city unseen test split), with \SI{1.7516}{dB} LoS and
\SI{3.3967}{dB} NLoS on the held-out cities. The spread priors achieved
\SI{28.10}{ns} delay-spread RMSE and \SI{13.74}{\degree} angular-spread RMSE
on the final 14-city unseen test split. The final residual GMM-head model
(retaining the GMM-head residual parameterisation) then improved those frozen
anchors on the unseen test cities, reaching \SI{1.6519}{dB} path-loss RMSE,
\SI{26.5570}{ns} delay-spread RMSE, and \SI{11.3854}{\degree}
angular-spread RMSE.
This means that the prior is not merely a weak baseline placed before the final
network. It is already a high-performing, train-calibrated estimator, and the
final model's contribution is measured against that demanding reference.

\subsection{What the final result means}

The path-loss result is the clearest success case. The path-loss prior already reached
\SI{1.9246}{dB} overall RMSE without a neural correction, which shows that a
large part of the CKM path-loss field is governed by stable radial and
visibility-dependent structure. The final model reduced this to \SI{1.6519}{dB}. This
improvement is numerically smaller than the early jumps between failed neural
models, but it matters more because it is obtained after the strong prior has
already removed the easy component of the task. The remaining error is mostly
the hard local residual: blocked streets, transition zones, and imperfect
geometry-to-radio alignment.

The spread targets should be interpreted differently. Delay spread and angular
spread are not smooth deterministic surfaces in the same way as LoS path loss:
they contain spike-like regimes, long tails, and topology-dependent mixtures.
For that reason, the final-model gains from \SI{28.10}{ns} to \SI{26.5570}{ns} in
delay spread and from \SI{13.74}{\degree} to \SI{11.3854}{\degree} in angular
spread are meaningful even though they are modest in absolute terms. They show
that the neural residual can add spatial context on top of the calibrated
distributional priors, but they also confirm that these targets remain less
deterministic than path loss.

The final model is therefore not a pure neural victory, and the prior is not
claimed as a revolutionary new propagation law. It is a division of labour.
Calibrated priors explain the part of the signal that is stable, low-variance,
and physically structured; the neural model is used where the prior is expected
to be incomplete. This is the main methodological result of the thesis:
performance improved when the model stopped trying to learn every effect from
pixels alone and instead learned bounded residuals around interpretable anchors.
In practical terms, the thesis validates a divide-and-conquer strategy.

\subsection{Answers to the research questions}

\textbf{RQ1: dense prediction under city holdout.} The project shows that dense
\(513\times513\) prediction is feasible under strict city-holdout evaluation,
but the successful formulation is not a plain U-Net-style regression problem.
Pure image-to-image models learned the smooth radial structure reasonably well
in LoS regions but repeatedly failed on NLoS tails and spread distributions.
The final priors satisfy the path-loss, delay-spread, and angular-spread error
targets, while the final model shows that a joint residual GMM-head framework can refine those
priors without discarding their physical meaning.

\textbf{RQ2: prior versus neural prediction.} The answer is target-dependent.
For LoS path loss, the calibrated enhanced two-ray prior is strong enough to
reach near-quantisation-scale errors, so forcing a neural model to rediscover
the same radial ring structure is inefficient. For NLoS path loss and the two
spread targets, distributional modelling and topology/regime information remain
essential. The final design therefore does not choose between physics and deep
learning: it freezes the part that is already explained by calibrated priors
and leaves the residual, spatially local correction to the network.

\textbf{RQ3: continuous UAV height.} The CKM dataset covers UAV antenna heights
from \SI{12}{m} to \SI{478}{m}. The height-bin analysis shows that error is not
uniform across this interval: the low-altitude, NLoS-heavy regime remains the
hardest, while high-altitude cases are often closer to the frozen prior. This
supports the use of explicit height conditioning and also explains why a single
headline RMSE is not enough to validate a UAV CKM model.

\subsection{Main contributions}

The first contribution is a reproducible evaluation contract for this dataset:
city-holdout splits, ground-only masking, LoS/NLoS reporting, topology
diagnostics, and UAV-height breakdowns. This matters because weaker random
splits or unmasked metrics can make a model look better without improving the
deployment-relevant task.

The second contribution is the distribution-first diagnosis. The late-stage
histograms showed that the previous PMHHNet family was not merely underpowered;
it was being asked to regress multi-modal and spike-plus-tail targets with
losses that reward a single central estimate. The distribution-first predecessors converted that
diagnosis into GMM-based dense predictors and demonstrated why marginal target
shape must guide the model head. The final validation histograms reported in
Chapter~\ref{chap:results} close that loop: the selected prior-anchored
residual GMM-head no longer shows the severe NLoS path-loss mode collapse,
while the spread statistics keep the remaining high-tail underestimation
visible.

The third contribution is the final prior family. The path-loss prior explains why the LoS
path-loss map has a strong radial structure and why an enhanced two-ray model
can outperform many neural experiments on that regime. The spread prior extends the same
principle to spread statistics with log-domain ridge calibration. Together,
these priors become interpretable anchors for the final joint residual GMM-head model
and a reusable baseline for later CKM experiments under the same dataset contract.

\subsection{Practical implications and scope}

The practical implication is that CKM prediction should be treated as a
model-selection problem, not only as a neural-architecture problem. The final
pipeline is useful because it identifies which part of the map can be predicted
with a stable prior and which part still needs learned spatial context. For
path loss, especially in LoS, the physically calibrated component is already
strong enough that a large CNN should not be asked to relearn it from scratch.
For delay spread, angular spread, and NLoS-heavy regions, the map is less
deterministic and the residual learner has a clearer role.

This also changes how future experiments should be judged. A new backbone is
only useful if it improves the held-out city metrics after the same masking,
height conditioning, and prior baselines are applied. Otherwise it may be
solving an easier problem or moving error between regimes. The thesis therefore
argues for reporting not only a single overall RMSE, but also LoS/NLoS splits,
height-bin behaviour, and topology-specific failures. These diagnostics made
the development path longer, but they prevented false progress.

The result is strongest as a reproducible surrogate for the CKM simulation
contract used in this work. It should not be read as a replacement for
ray-tracing in arbitrary cities, frequencies, or antenna configurations.
Instead, it is evidence that a simulation-derived CKM can be compressed into a
fast predictor when the dataset contract is fixed and the predictor is anchored
to interpretable propagation structure. This is a narrower claim, but it is a
more defensible one.

The final Barcelona runtime benchmark supports the acceleration claim under a
like-for-like full-resolution setup. Pixel-level MATLAB ray tracing took
\SI{99.89}{s} per CKM sample, while the complete final-generator path took
\SI{0.88}{s} on the same local laptop, including geometric LoS/NLoS mask
generation, the final calibrated path-loss and spread priors, and the PL/DS/AS
neural prediction. The
resulting \(113.7\times\) reduction exceeds the original 10 to 100\(\times\)
runtime goal. It uses a vectorized CUDA LoS/NLoS backend that
was checked to produce zero difference with the original GT ray-caster on the
Barcelona and exported CKM samples, plus a torch implementation of the frozen
calibrated priors. After these optimizations, the mask step takes \SI{0.30}{s},
the priors \SI{0.04}{s}, and the mixed-precision residual GMM-head inference
itself about \SI{0.54}{s} per sample.

For deployment-oriented work, the main value of the pipeline is transparency.
The priors expose the physical assumptions, the residual caps bound how much
the network can override them, and the evaluation protocol reveals where the
model remains weak. That combination is more useful for UAV network planning
than a black-box map generator with a slightly lower aggregate error but no
clear indication of when it is extrapolating.

\subsection{Negative results}

Several abandoned directions were scientifically useful. The pix2pix-style cGAN
line produced sharper-looking maps but did not improve the physical RMSE that
matters for propagation. Six-way topology experts and PMHHNet variants improved
organisation and reduced some overfitting, but they did not remove the NLoS
mode-collapse problem. Multi-scale and quad-head experiments added supervision at
several resolutions, yet they did not solve the deeper distribution mismatch.
The heteroscedastic uncertainty diagnostic showed that hard pixels can be identified, but it
was not enough by itself to become the final predictor.

These failures are not side notes. They are the reason the thesis ends with a
hybrid design rather than with a larger black-box CNN. Each negative result
removed an explanation that was too simple: the problem was not only decoder
artifacts, not only missing physical channels, not only regularisation, and not
only lack of capacity. The final approach is more convincing because it is built
on those rejected hypotheses.

\subsection{Lessons learned across the development path}

Table~\ref{tab:conclusion_lessons} summarises the main lesson from each branch
of the project. The sequence matters because the final pipeline is not an
isolated model choice; it is the result of progressively narrowing down what
the dataset would and would not reward.

\begin{table}[htbp]
\centering
\caption{Main lessons learned across the development path.}
\label{tab:conclusion_lessons}
\scriptsize
\renewcommand{\arraystretch}{1.15}
\begin{tabularx}{\textwidth}{>{\raggedright\arraybackslash}p{0.20\textwidth}
>{\raggedright\arraybackslash}X
>{\raggedright\arraybackslash}X}
\toprule
Development branch & What it showed & Consequence for the final thesis \\
\midrule
U-Net and cGAN baselines &
Dense image-to-image models can reproduce smooth map structure, but visual
sharpness is not the same as propagation accuracy. &
The thesis kept dense prediction, but stopped treating perceptual map quality
as the main objective. \\
Masked baselines &
Building pixels can distort both loss and metrics if they are treated as normal
prediction targets. &
Ground-only training and reporting became part of the evaluation contract. \\
PMNet and PMHHNet variants &
More specialised CNN structure improved organisation but did not solve the
multi-modal NLoS and spread-target errors. &
The project shifted from architecture search toward target-distribution
diagnosis. \\
Distribution-first models &
GMM and heteroscedastic heads exposed spike-plus-tail behaviour and hard-pixel
uncertainty. &
The spread targets were treated as distributional, not simply as smooth scalar
regression maps. \\
Final priors &
Calibrated physical and statistical priors could beat many neural experiments on
their intended regimes. &
The final system used priors as frozen anchors rather than as optional
features. \\
Final residual GMM-head model &
Residual learning improved the frozen anchors without erasing their
interpretability. &
The final claim became a hybrid one: calibrated priors plus bounded neural
correction under strict city holdout. \\
\bottomrule
\end{tabularx}
\end{table}

\section{Future Directions}

\paragraph{Short-medium term.}
The next step is to strengthen the current residual pipeline without changing
the evaluation contract, then extend it toward frequency and radio-configuration
conditioning. Useful additions include transmitter-oriented mixup, explicit
corridor losses, and controlled ablations of each residual head and loss
weight. The current
dataset fixes the carrier and antenna assumptions, while a more general outdoor
CKM model should learn how path loss and spreads vary with frequency,
bandwidth, antenna pattern, and possibly polarization. A natural dataset for
this stage would simulate the same cities across several radio configurations
so that the model can separate geometry effects from configuration effects.

A nearer-term training direction is adaptation to new or under-represented
topology types. The per-topology analysis suggests that dense high-rise,
compact mid-rise, open low-rise, mixed urban layouts, and future morphology
classes do not fail in exactly the same way. A practical continuation would
therefore be to fine-tune the final prior-anchored model on newly introduced
urban typologies, or to train lightweight specialist heads on top of the shared
backbone when only limited examples are available. This would test whether the
model can recover local structure in topology regimes beyond the original
training distribution, rather than only specialising to already defined classes.

For delay and angular spreads, another short- to medium-term direction is to
optimize not only pixel-wise RMSE but also spatial structure. Spread maps can
look physically more coherent even when small shifts across quantized target
boundaries are penalized harshly by RMSE. Future fine-tuning should therefore
test structural objectives, for example gradient consistency, region-wise
correlation, SSIM-like terms, LoS/NLoS boundary consistency, or topology-aware
ranking losses, so that the model preserves the correct spread patterns rather
than only minimizing the average numeric error.

\paragraph{Medium-long term.}
The final extension is field calibration. The path-loss and spread priors are
interpretable, which makes them suitable for simulator-to-reality correction
using a limited number of real measurements. In the medium-long term, this could lead
to a CKM predictor that starts from ray-traced or prior-generated maps, adapts
to a new city with sparse measurements, and still reports uncertainty by
LoS/NLoS regime, UAV-height bin, and topology class. Those safeguards should
remain part of the method before considering operational use in emergency
coverage, UAV network planning, or rapid deployment studies.
