This section provides the technical documentation for the non-DL physical
priors (\texttt{Try 78} and \texttt{Try 79}) and the final joint architecture
(\texttt{Try 80}). It serves as a bridge between the high-level attempt history
and the numerical results, explaining the exact mathematical transformations,
how every parameter was fitted in practice, and what is genuinely new in the
final \texttt{Try 80} approach.

\paragraph{Notation convention.}
Following the notation used in the ETSETB template, vectors and matrices are
typeset in bold roman font. Vectors use lowercase symbols and matrices, maps or
tensors use uppercase symbols. Scalars are left in italic math font. For
example, \(\textbf{x}_n\) denotes the feature vector of pixel \(n\),
\(\textbf{X}\) denotes a design matrix, \(\textbf{w}\) denotes a vector of
fitted coefficients, and \(y_n\), \(d_{2D,n}\), or \(h_{\mathrm{tx}}\) denote
scalar quantities. The implementation often stores full maps as tensors, but
the derivations below use a pixel index \(n\) whenever this makes the
regression or likelihood clearer.

\begin{table}[H]
\centering
\small
\begin{tabularx}{\textwidth}{p{2.2cm} p{3.1cm} p{4.0cm} X}
\toprule
\textbf{Block} & \textbf{Literature basis} & \textbf{CKM adaptation} &
\textbf{Role in \texttt{Try 80}} \\
\midrule
Path-loss LoS & FSPL and coherent two-ray propagation &
Effective height-dependent reflection parameters and radial residuals fitted
from CKM LoS pixels &
Frozen LoS prior, injected as an input channel and used as the anchor for
bounded residual learning. \\
\addlinespace
Path-loss NLoS & COST-231/Hata-style urban loss and A2G elevation envelopes &
Regime-wise calibration with local density, height, shadowing and elevation
features &
Frozen NLoS prior selected by the explicit LoS/NLoS geometry mask. \\
\addlinespace
Delay and angular spread & Log-domain large-scale spread modelling and
LoS/NLoS separation &
Regime-wise ridge regression in \(\log(1+y)\) space using multiscale
morphology and obstruction features &
Frozen spread priors used as anchors for the multi-task residual model. \\
\addlinespace
\texttt{Try 80} residual model & Probabilistic regression and mixture-density
ideas &
Prior-anchored per-pixel GMM with residual caps, anchor gates, explicit
region masks and shared multi-task trunk &
Learns residual uncertainty and local corrections without replacing the
interpretable priors. \\
\bottomrule
\end{tabularx}
\caption{One-page summary of what is inherited from the literature, what is
adapted to CKM, and what is reused or newly assembled in \texttt{Try 80}.}
\label{tab:executive_summary}
\end{table}

\subsection{High-level architecture}

The central design decision in \texttt{Try 80} is \emph{not} to learn path
loss, delay spread, and angular spread from scratch. Instead:

\begin{itemize}
\item \texttt{Try 78} provides a frozen prior for \texttt{path\_loss},
\item \texttt{Try 79} provides frozen priors for \texttt{delay\_spread} and
      \texttt{angular\_spread},
\item \texttt{Try 80} learns bounded residual corrections around those priors.
\end{itemize}

The deterministic or interpretable structure is kept in the prior, and the deep
model is used only for the residual part that the prior cannot explain.
Figure~\ref{fig:pipeline} illustrates the overall pipeline.

\begin{figure}[H]
\centering
\begin{tikzpicture}[
  box/.style={rectangle, draw, rounded corners=3pt, minimum width=2.2cm,
              minimum height=0.75cm, align=center, font=\footnotesize},
  arr/.style={-{Stealth[length=5pt]}, thick},
  node distance=0.5cm and 1.1cm
]
\node[box, fill=blue!15]  (t78) {Try 78\\(path loss)};
\node[box, fill=green!15, below=of t78] (t79) {Try 79\\(spreads)};
\node[box, fill=orange!15, right=1.2cm of t78, yshift=-0.62cm] (inject)
      {Prior injection\\(input channels)};
\node[box, fill=red!15, right=1.2cm of inject] (dnn)
      {DNN residual\\multi-task};
\node[box, fill=gray!20, right=1.2cm of dnn] (out)
      {PL, DS, AS\\outputs};

\draw[arr] (t78) -- (inject);
\draw[arr] (t79) -- (inject);
\draw[arr] (inject) -- (dnn);
\draw[arr] (dnn) -- (out);

\node[below=0.25cm of dnn, font=\scriptsize, text width=2.4cm, align=center]
      {$\delta \in [-\delta_{\max}, +\delta_{\max}]$};
\end{tikzpicture}
\caption{Overall \texttt{Try 80} pipeline. The frozen priors are injected into the DNN, which learns a bounded correction $\delta$ around them.}
\label{fig:pipeline}
\end{figure}

\subsection{How to construct a good prior}
\label{sec:good_prior_design}

The most important lesson from the prior phase is that a useful prior is not
just a formula that looks physical. A good prior is a complete, internally
consistent measurement-to-prediction pipeline. It must use the same formula at
fitting time and inference time, compute its features in the same units in both
places, respect the evaluation mask, and encode only those physical assumptions
that are actually visible in the data. This distinction explains why the
earlier prior attempts around \texttt{Try 41}--\texttt{Try 50} looked very
similar to the final \texttt{Try 78} prior, but produced radically different
NLoS results.

\subsubsection{A prior is more than an analytical equation}

The first physical-prior experiments used the correct high-level idea:
\begin{equation}
  \widehat{\mathrm{PL}}(x)
  =
  \mathrm{PL}_{\mathrm{prior}}(x)
  +
  r_{\theta}(x),
  \label{eq:good_prior_residual}
\end{equation}
where the analytical map explains the large-scale propagation trend and the
network or calibration only predicts what remains. This idea was already much
better than asking a CNN to rediscover path loss from raw topology alone.
However, the numerical result showed that the first prior pipeline was still
not good enough: the raw physical formula gave approximately
\SI{67.23}{dB} RMSE, train-only regime-aware calibration improved it to about
\SI{24.16}{dB}, and the enhanced obstruction-aware prior used around
\texttt{Try 45} still reported approximately \SI{41.27}{dB} NLoS RMSE.

The mistake would be to interpret this as evidence that physical priors were
useless. The later results show the opposite. The real issue was that the
prior was being used as a \emph{collection of formulas}, while the final
\texttt{Try 78} prior treats it as a reproducible estimator. A formula can be
correct in isolation and still fail if the fitted calibration is attached to a
different runtime formula, or if the morphology features used during fitting
are scaled differently from those used during evaluation. This is exactly the
kind of failure that hurts NLoS most, because NLoS is controlled by local
obstructions, nearby building density, elevation angle, and regime identity
rather than by distance alone.

\subsubsection{Why the old prior and \texttt{Try 78} were not equivalent}

The earlier prior line and \texttt{Try 78} shared many ingredients:
FSPL/two-ray geometry, COST-231-style urban attenuation, LoS/NLoS regime
separation, antenna-height bins, and local obstruction features. On a slide,
they can look almost identical. The difference is that \texttt{Try 78} removed
several hidden inconsistencies.

First, the base formula and the fitted calibration were made to match. In the
old \texttt{Try 47}-derived prior, the calibration had been fitted for an
A2G-strengthened prior of the form
\(\mathrm{hybrid\_two\_ray\_cost231\_a2g\_nlos}\), but some runtime paths used
the simpler \(\mathrm{hybrid\_two\_ray\_cost231}\) prior channel. The fitted
coefficients therefore corrected one prior while inference sometimes supplied
another. This is a severe error for NLoS pixels, because the A2G-NLoS envelope
is precisely the part that changes the blocked-region baseline.

Second, \texttt{Try 78} removed a feature-scale mismatch. In the older
pipeline, the calibration script and the runtime data path did not compute the
local-height features in exactly the same scale. The fitting code saw height
features derived from raw or near-raw topology values, whereas the runtime path
could apply the calibration to already normalized tensors and then divide by a
height scale again. This made the same physical quantity appear with different
magnitudes during fitting and inference. The calibration coefficients then had
to compensate for an implementation mismatch instead of learning a stable
physical correction. Because NLoS loss depends strongly on local building
height and obstruction morphology, this mismatch was much more damaging in
NLoS than in LoS.

Third, regime assignment became sample-based instead of mainly city-label
based. The earlier calibration inherited city-type mappings from previous
experiments. This is useful when the split is stable, but it is less robust
when a specific sample differs from the average morphology of its city.
\texttt{Try 78} derives the coarse regime from the sample topology itself:
building density, mean building height, LoS/NLoS mask, and antenna-height bin.
This makes the prior closer to the actual input geometry and better aligned
with city-holdout generalization.

Fourth, the whole prior was evaluated as a self-contained prior-only system.
\texttt{Try 78} loads the HDF5 arrays directly, computes the propagation
formula, constructs the feature vector, fits the calibration on fit cities, and
evaluates on held-out cities. This removed the ambiguity introduced by a larger
training pipeline with normalized tensors, YAML-controlled formula channels,
and separate calibration scripts. The result was not just cleaner code; it was
a different scientific object. It became possible to say exactly what the
prior knows and exactly how it is fitted.

\subsubsection{Scope of this mismatch}

It is important not to overgeneralize this diagnosis. The feature-scale
mismatch described above was a real problem in the \emph{prior calibration}
path, especially in the branch descended from \texttt{Try 47}. It does not
mean that all bad NLoS results in earlier deep-learning branches were caused
by the same hidden normalization bug.

The no-prior and weak-prior DL branches had other verified problems. Some
late residual experiments disabled or weakened the main full-resolution loss;
some used excessive regularization, causing predictions to remain too close to
the prior; some used Huber or smoothed objectives that reduced sensitivity to
large NLoS tail errors; and some corridor-weighting settings over-emphasized
the centre of the map. These were training-design issues, not the same
fit-vs-runtime calibration-scale mismatch.

The clearest evidence is that the bad NLoS behaviour appears even in branches
where the prior path was absent or disabled. The clean \texttt{Try 44}
no-prior PMNet-v3 control learned the LoS part reasonably well
(\(\mathrm{RMSE}_{\mathrm{LoS}}\approx\SI{4.27}{dB}\)), but still failed in
NLoS (\(\mathrm{RMSE}_{\mathrm{NLoS}}\approx\SI{38.58}{dB}\)). This run did
not depend on the obstruction-calibrated formula prior, so its NLoS error
cannot be explained by the prior's fit-vs-runtime feature mismatch. Later
no-prior reruns showed the same pattern: for example, \texttt{Try 60}
reported about \SI{3.94}{dB} LoS RMSE and \SI{50.50}{dB} NLoS RMSE on the
open-sparse vertical expert; adding stronger NLoS supervision in
\texttt{Try 61} improved the lowrise NLoS number to about \SI{36.30}{dB}, but
did not solve it. \texttt{Try 73} explicitly disabled the formula-prior input,
the obstruction-feature input, and the residual-over-prior objective
(\texttt{enabled = false} for the corresponding configuration blocks);
nevertheless, its
no-prior experts still showed NLoS errors around
\(\SI{41}{dB}\)--\(\SI{49}{dB}\) in the available validation summaries, while
LoS stayed near \SI{4}{dB}--\SI{6}{dB}. In other words, the network could learn
the smooth radial/free-space structure, but it could not learn the blocked
NLoS distribution from direct point regression alone.

\texttt{Try 76} was therefore not just a cleanup of the prior pipeline. It was
built from scratch without the physical prior and diagnosed the deeper failure
mode: the old DL family tended toward NLoS mode collapse, with predicted NLoS
means around \SI{36.5}{dB} against target means around \SI{107.1}{dB}. That
pattern is not the signature of a small feature-scale bug. It is the signature
of a point-regression model being structurally mismatched to the NLoS target
distribution. The histogram study showed that NLoS path loss is better
described by a multi-component distribution than by a single Gaussian-like
output, and that large NLoS errors are rare enough to be underweighted by
ordinary pixel losses but large enough to dominate RMSE.

Therefore the correct interpretation is twofold. The old prior branch really
was harmed by formula and feature-scaling inconsistencies, and fixing those
inconsistencies explains why \texttt{Try 78} became so strong. However, the
bad NLoS behavior of earlier direct DL models should not be dismissed as a
prior-calibration bug. Those models were also intrinsically weak for NLoS
because their losses and output distributions did not represent the blocked,
multi-modal, heavy-tailed regime well enough. This is why the thesis keeps
\texttt{Try 76} as a genuine modelling advance, not merely as a cleanup of the
prior pipeline.

\subsubsection{What changed in the LoS branch}

The final \texttt{Try 78} LoS prior is also not merely the old
two-ray/COST-ish branch with better bookkeeping. The early physical priors used
an incoherent or piecewise large-scale two-ray idea: FSPL at shorter ranges and
a plane-earth two-ray trend at longer ranges. This captures the fact that
clear propagation is radial and height-dependent, but it does not model the
ring-like interference pattern visible in the CKM LoS maps.

The final LoS branch instead uses a fitted coherent two-ray model. The received
field is treated as a complex sum of a direct path and an image-source
reflected path. Rather than assuming a textbook Fresnel reflection coefficient
for a perfectly flat material, the model fits an effective complex reflection
coefficient
\[
  \Gamma(h_{\mathrm{tx}})
  =
  \rho(h_{\mathrm{tx}})\,e^{j\phi(h_{\mathrm{tx}})}
\]
per UAV-height bin. This gives the model enough freedom to represent the CKM
simulator's effective ground-reflection behaviour without pretending that the
urban scene is a perfectly smooth dielectric plane. A height-dependent bias
and a small radial residual correction are then added to remove systematic
range-dependent errors.

This is the reason for the large LoS jump. FSPL alone reached about
\SI{3.9540}{dB} LoS RMSE. Adding a simple radial residual improved only
slightly, to about \SI{3.7086}{dB}. The coherent two-ray/ring model reduced
LoS RMSE to about \SI{1.7516}{dB}. In other words, the dominant LoS error was
not random urban complexity. It was a deterministic radial interference
pattern that the old prior family had not represented explicitly. Once that
pattern was modelled as coherent interference, the LoS branch became almost a
closed-form reconstruction of the simulator's large-scale behaviour.

\subsubsection{What changed in the NLoS branch}

The NLoS improvement is more subtle. \texttt{Try 78} did not discover a
single magic NLoS formula. It made the old obstruction-aware idea usable by
making every component agree with the others. The raw NLoS branch combines a
COST-231/Hata-style urban attenuation curve with an elevation-aware A2G
envelope and keeps the more conservative prediction:
\[
  \mathrm{PL}_{\mathrm{NLoS,raw}}(x)
  =
  \max
  \left(
    \mathrm{PL}_{\mathrm{COST231}}(x),
    \mathrm{PL}_{\mathrm{A2G,NLoS}}(x)
  \right).
\]
This is still only a baseline. It says that low-elevation, obstructed links
should be more lossy than clear links, but it cannot know whether a particular
street pixel lies in a shallow transition region, a deep shadow, or a dense
urban canyon.

The decisive step is the calibrated feature vector. For each pixel,
\texttt{Try 78} constructs features from the prior value, log-distance,
multiscale building density, multiscale building height, multiscale NLoS
support, elevation angle, and an empirical shadow-variance proxy. These
features are then fitted separately by coarse topology class, LoS/NLoS state,
and antenna-height bin. The prior therefore does not use one global correction
for all blocked pixels. It learns a different correction for, for example, a
low-altitude NLoS pixel in a dense high-rise map than for a high-altitude NLoS
pixel in an open low-rise map.

This design is close in spirit to the literature, but not identical to it.
The literature provides the physical axes: distance, frequency, height,
elevation angle, LoS/NLoS state, and scenario class. Our contribution is to
turn the CKM topology map into the missing local scenario descriptors. Instead
of assigning the whole map to one scenario and applying a single empirical
curve, the model computes local density and obstruction support around every
receiver pixel. That is what makes the prior map-aware rather than just
scenario-aware.

\subsubsection{What was improved relative to the literature}

The final prior should be read as a hybrid between classical propagation
models and dataset-specific calibration. The classical part is important
because it gives the correct coordinate system: path loss is primarily a
function of distance, frequency, antenna height, elevation angle, and
propagation state. The dataset-specific part is equally important because CKM
is not a textbook macrocell measurement campaign. It is a dense urban
ray-tracing map with a centred UAV transmitter, a fixed receiver height, a
\SI{7.125}{GHz} carrier, and a full topology raster for every sample.

Compared with applying literature formulas directly, the final prior improves
four things.

\begin{enumerate}
\item \textbf{It replaces assumed ground material with fitted effective
      reflection.} A textbook two-ray model would require a dielectric
      constant, conductivity, polarization, and a flat ground assumption.
      \texttt{Try 78} instead fits \(\rho\), \(\phi\), and bias from CKM LoS
      pixels, which captures the effective simulator behaviour.
\item \textbf{It keeps the coherent interference pattern.} Many simplified
      two-ray uses collapse the model into a monotonic distance law. The CKM
      LoS maps contain rings, so preserving the complex field sum is essential.
\item \textbf{It localizes the scenario description.} 3GPP-style models define
      broad scenarios such as urban macro or urban micro. The prior keeps that
      idea but computes local morphology features directly from the topology
      map, at 15 m and 41 m scales.
\item \textbf{It separates physical extrapolation from empirical correction.}
      FSPL, two-ray, COST-231, and A2G envelopes define the shape of the prior;
      train-only calibration adapts that shape to CKM without letting the
      calibration see the evaluation cities.
\end{enumerate}

The practical lesson is that a good prior is neither a pure textbook formula
nor a black-box regressor. A pure formula is too rigid for dense CKM maps, and
a black-box regressor can lose the physical coordinate system. The successful
prior is in the middle: it starts from literature-backed propagation structure,
then uses the dataset to fit the parameters and local correction terms that the
literature cannot know.

\subsubsection{Resulting design rules}

From the failed and successful prior attempts, the following rules were used
for the final \texttt{Try 78}/\texttt{Try 80} construction:

\begin{itemize}
\item \textbf{Use the prior only on valid receiver pixels.} Fitting and
      evaluation must use the same ground-pixel mask as the final metric.
\item \textbf{Fit and apply the same formula.} The formula used to generate
      training features must be exactly the formula used at inference time.
\item \textbf{Keep units explicit.} Distances, heights, topology values, and
      normalized features must have one definition shared by fitting and
      inference.
\item \textbf{Hard-separate LoS and NLoS when the dataset provides a reliable
      mask.} The two regimes have different physics and should not be forced
      through one global curve.
\item \textbf{Let the literature define the axes, not all coefficients.}
      Distance, height, frequency, and elevation angle come from propagation
      theory; effective reflection and morphology corrections are fitted from
      CKM.
\item \textbf{Prefer bounded residual learning after the prior.} Once the
      prior explains the deterministic structure, the neural model should make
      local corrections instead of being allowed to overwrite the prior
      arbitrarily.
\end{itemize}

These rules explain why \texttt{Try 78} could reach
\SI{1.9246}{dB} overall path-loss RMSE, with \SI{1.7516}{dB} LoS and
\SI{3.3967}{dB} NLoS, while earlier priors with similar names and similar
ingredients were still stuck around \SI{40}{dB} NLoS. The improvement came
from using the right physical structure \emph{and} making the complete
calibration pipeline internally consistent.

\subsection{What was reused from the literature, what was changed, and what is new}

The \texttt{Try 80} approach is not a single literature formula, but a
multi-stage construction. The following points summarize the technical
decisions for each task.

\subsubsection{Try 78: path loss}

\paragraph{What the literature provides.}
FSPL is a textbook formula \cite{rappaport} that gives the power loss of a
plane wave in free space as a function of distance and frequency. The two-ray
model for a transmitter above a flat reflecting ground is also classical
\cite{rappaport}. The COST-231 extension of the Hata model is a standard
empirical formula for macrocell path loss in the 1.5--2 GHz range.
Treating LoS and NLoS as fundamentally different propagation regimes is
standard practice in 3GPP TR 38.901 \cite{tr38901}.

\paragraph{What we took exactly from the literature.}
The FSPL formula is used verbatim. The two-ray concept---direct ray plus
ground-reflected ray with a complex-valued sum---is taken directly from the
standard derivation in \cite{rappaport}. The COST-231 formula is used
verbatim including all empirical constants. The qualitative idea of LoS/NLoS
regime separation is taken from \cite{tr38901}.

\paragraph{What we changed or did not take.}
\begin{itemize}
\item \textbf{Reflection coefficient $\Gamma$.} The coherent two-ray equation
      itself is kept, but the material-based Fresnel coefficient is replaced by
      an \emph{effective} complex parameter
      $\rho(h_{\mathrm{tx}}) e^{j\phi(h_{\mathrm{tx}})}$ fitted from data.
\item \textbf{Height-bin fitting.} Independent $(\rho, \phi, b)$ values are
      fitted per 5~m UAV-height bin and linearly interpolated between
      neighboring bins at inference, producing a height-dependent effective
      model needed because the CKM dataset covers roughly 10--460 m of UAV
      height.
\item \textbf{Radial residual correction.} A separate range-dependent
      correction term $r(d_{2D})$ is added after the standard two-ray
      prediction to capture systematic smooth trends in its residuals.
\item \textbf{Regime-wise linear calibration on top of COST-231.} A 15-feature
      design vector $\textbf{x}_{78}(x)$ and regime-wise OLS calibration correct the
      raw prior using local density, height, NLoS support, and elevation
      features. This step has no direct literature equivalent.
\end{itemize}

\subsubsection{Try 79: delay spread and angular spread}

\paragraph{What the literature provides.}
3GPP TR 38.901 \cite{tr38901} and WINNER II \cite{winner2} model delay spread
and angular spread as log-normal random variables. Separate parameters are
given for LoS and NLoS, and for different scenario types. Empirical UAV studies
\cite{cai2019,izydorczyk2019} report that delay and angular spread increase as
the UAV elevation angle decreases, which was corroborated by inspecting the
dataset data.

\paragraph{What we changed or did not take.}
\begin{itemize}
\item \textbf{Log-domain modelling.} We reuse the literature assumption
      that spreads are log-normally distributed, so we predict $\log(1+y)$.
\item \textbf{Morphology features.} Standard models use distance and antenna
      height; we extend this with local building density and height features
      extracted from the CKM map.
\end{itemize}

\subsubsection{Try 80: architecture and loss}

\begin{itemize}
\item \textbf{No stochastic channel simulator.} We fit a deterministic
      regression that predicts the \emph{expected} spread value at each map
      pixel, conditioned on local geometry.
\item \textbf{Map-derived morphology features.} Per-pixel features such as
      local building density $\rho_k(x)$, mean building height $h_k(x)$,
      NLoS-support fraction $n_k(x)$, blocker depth $b_{41}(x)$, transmitter
      clearance $c_{41}(x)$, and taller-building fraction $t_{41}(x)$ have
      no equivalent in 3GPP or WINNER.
\item \textbf{Regime-wise ridge regression per pixel.} The regime-wise ridge
      regression and the associated fallback hierarchy are entirely our
      construction, designed to handle the sparse-regime problem that arises
      when the dataset is stratified across
      $\mathrm{metric} \times \mathrm{topology} \times \mathrm{LoS/NLoS}
      \times \mathrm{antenna\_bin}$.
\item \textbf{Shared UNet with FiLM.} The backbone is a standard medical/vision
      U-Net, but with sinusoidal FiLM conditioning on UAV height.
\item \textbf{Prior-anchoring.} This is a genuinely new design choice for
      this project: the model can only learn small corrections $[-\delta,
      +\delta]$ relative to the physical priors.
\item \textbf{UAV height as a feature.} The normalized height
      $h_{\mathrm{norm}}$ and its square $h_{\mathrm{norm}}^2$ are included
      because the CKM dataset spans a wide range of UAV heights.
\end{itemize}
